---
title: "Co-flowering and hummingbird overlap in Neotropical Gesneriaceae: Phylogeny and Interactions"
author: "Ben Weinstein"
date: "5/30/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r,message=F,warning=F}
library(knitr)
opts_chunk$set(message=F,warning=F,echo=F)

library(ape)
library(kableExtra)
library(boot)
library(dplyr)
library(ggplot2)
library(phytools)
library(stringr)
library(splines)
library(R2jags)
library(reshape2)
library(tidyr)
```

# Interactions

```{r}
#Interactions
interactions<-read.csv("data/cleaned/interactions.csv",row.names=1)
head(interactions) %>% kable() %>% kable_styling()
```

# Phylogeny

```{r}
phy<-read.newick("data/SantaLucia_tree_Newick")
species<-str_match(phy$tip.label,"_(\\w+)")[,2]
tspecies<-unique(interactions$Plant)
to_join<-data.frame(Plant=tspecies,phylo=str_match(tspecies," (\\w+)")[,2])
new_names<-data.frame(phylo=species) %>% inner_join(to_join)
phy$tip.label<- as.character(new_names$Plant)
plot(phy)

#variance covariance matrix
vcov<-vcv(phy,corr = T)

#Identity matrix
I<-diag(nrow(vcov))
```

# Traits
```{r}
traits<-read.csv("data/Species_SantaLucia.csv")

#improve column names
traits<-traits %>% select(Plant=Iplant_Double,min_corolla_tube=Lt,corolla_length=Lc,vertical_diameter= D,stamen_length=S,stamen_exertion=E)

#order the factors

head(traits) %>% kable() %>% kable_styling()
```

Sum interaction matrix plotted in plant phylogenetic order.

```{r}
m<-interactions %>% filter(Plant %in% phy$tip.label) %>% group_by(Hummingbird,Plant) %>% summarize(n=n())
m$Plant<-factor(m$Plant,levels=phy$tip.label)
ggplot(m,aes(x=Hummingbird,y=Plant,fill=n)) + geom_tile() + theme_bw() + theme(axis.text.x=element_text(angle=-90)) + scale_fill_continuous(low="blue",high="Red")
```

# Estimate interaction probability

The goal is to generate a distance matrix of mean niche overlap among plants.

```{r}
ggplot(m,aes(x=Hummingbird,y=Plant,fill=n>0)) + geom_tile() + theme_bw() + theme(axis.text.x=element_text(angle=-90))
```

```{r}
#Impute non-detections. Start by assuming all hummingbirds occur at all elevations (can be relaxed)

sumf<-interactions %>% group_by(Hummingbird,Plant,ID,Date,ele) %>% dplyr::summarize(Yobs=n())

#Create emptys zeros
impute_zero<-function(x,Hummingbird){
  df<-data.frame(Hummingbird=Hummingbird)
}

zero_count<-sumf %>% group_by(Plant,ID,Date,ele) %>% do(impute_zero(.,Hummingbird=levels(sumf$Hummingbird)))
zero_count<-zero_count %>% anti_join(sumf)

#add a dummy label for presence absence
indat<-bind_rows(sumf,data.frame(zero_count,Yobs=0))
```

```{r}

#align factors with phylogeny
indat<-indat %>% filter(Plant %in% phy$tip.label)

#Easiest to work with jags as numeric ordinal values
indat$Hummingbird<-as.factor(indat$Hummingbird)
indat$Plant<-factor(indat$Plant,levels=phy$tip.label)
indat$jBird<-as.numeric(indat$Hummingbird)
indat$jPlant<-as.numeric(indat$Plant)

jagsIndexBird<-data.frame(Hummingbird=levels(indat$Hummingbird),jBird=factor(1:length(levels(indat$Hummingbird))))
jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=factor(1:length(levels(indat$Plant))))

#Turn Time and ID into numeric indexes
indat$jTime<-as.numeric(as.factor(indat$Date))
indat$jID<-as.numeric(as.factor(indat$ID))
```

```{r}
#hold model objects
models<-list()
chains<-list()
```

# Poisson Interactions

```{r}
#Source model
source("model/poisson_interactions_detection.R")

#print model
writeLines(readLines("model/poisson_interactions_detection.R"))

#Run Model
runModel<-function(Yobs_dat){
  
  #Parameters to track
  ParsStage <- c("alpha","omega")
  
  InitStage<-function(){
    list(z=array(dim=c(Dat$Birds,Dat$Plants,Dat$Cameras),data=max(Dat$Yobs)))
  }
  #Jags Data

  Dat<-list(
    Yobs=Yobs_dat$Yobs,
    Birds=max(c(Yobs_dat$jBird)),
    Bird=Yobs_dat$jBird,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    Cameras=max(Yobs_dat$jID),
    Camera=Yobs_dat$jID,
    Nobs=length(Yobs_dat$Yobs))
  
  #MCMC options
    system.time(
      model<-jags(data=Dat, parameters.to.save=ParsStage,model.file="model/poisson_interactions_detection.jags",n.thin=1,inits = InitStage,n.iter=2000,n.burnin=1800,n.chains=2,DIC=F)
    )
    return(model)
}

models$intercept<-runModel(Yobs_dat=indat)
```

View Chains

```{r}
getChains<-function(mod){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
    #single index
  splitpc[c("omega")]<-lapply(
    splitpc[c("omega")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)"))[,3]
    pc<-data.frame(x,jBird=sv)
    })
  
  #double index
  splitpc[c("alpha")]<-lapply(
    splitpc[c("alpha")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+)]"))[,3:4]
    colnames(sv)<-c("jBird","jPlant")
    pc<-data.frame(x,sv)
    })
  
  chains<-bind_rows(splitpc)
return(chains)}
```

```{r,fig.height=8,fig.width=12}
chains$intercept<-getChains(models$intercept)

#merge species names
chains$intercept<-chains$intercept %>% left_join(jagsIndexBird) %>% left_join(jagsIndexPlants)

#alpha
chains$intercept %>% filter(parameter %in% "alpha",jPlant %in% c(5,10)) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~Hummingbird,scales="free")

chains$intercept %>% filter(parameter %in% "omega") %>% ggplot(.,aes(x=Hummingbird,y=value)) + geom_boxplot() + coord_flip()

```

### P(Interaction)

```{r,fig.height=8,fig.width=8}
chains$intercept %>% filter(parameter %in% "alpha") %>% group_by(Hummingbird,Plant) %>% summarize(n=mean(value)) %>% ggplot(.,aes(fill=n,x=Plant,y=Hummingbird)) + geom_tile() + theme(axis.text.x = element_text(angle=-90)) + labs(fill="P(Interaction)") + scale_fill_continuous(low="blue",high="Red")
```

# Niche Overlap

Euclidean Distance in the P(Interaction)

```{r}
alpha_matrix<-chains$intercept %>% filter(parameter %in% "alpha") %>% group_by(Hummingbird,Plant) %>% summarize(n=mean(value)) %>% acast(Hummingbird~Plant)
write.csv(alpha_matrix,"PInteraction.csv")
```

#Phylogenetic model of species interactions

Do plants show phylogenetic signal in hummingbird partners?

Start with binary interactions

```{r}
#Source model
source("model/intercept_phylogeny.R")

#print model
writeLines(readLines("model/intercept_phylogeny.R"))

#Run Model
runModel<-function(Yobs_dat,I,vcov){
  
  
  #Parameters to track
  ParsStage <- c("alpha","lambda","sigma","e","tauC")
  
  #Jags Data


  Dat<-list(
    Yobs=Yobs_dat$Yobs,
    Birds=max(c(Yobs_dat$jBird)),
    Bird=Yobs_dat$jBird,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    Nobs=length(Yobs_dat$Yobs),
    I=I,
    vcov=vcov,
    zeros=rep(0,max(c(Yobs_dat$jPlant))))
  
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/intercept_phylogeny.jags",n.thin=1,n.iter=100,n.burnin=0,n.chains=2,DIC=F)
    )
    return(model)
}

models$intercept<-runModel(Yobs_dat=indat,I=I,vcov=vcov)
```

#Get Chains
```{r}
getChains<-function(mod){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("alpha","tauC")]<-lapply(
    splitpc[c("alpha","tauC")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+)]"))[,3:4]
    colnames(sv)<-c("jBird","jPlant")
    pc<-data.frame(x,sv)
    })
  
    splitpc[c("e")]<-lapply(
    splitpc[c("e")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=sv)
    })
    
  chains<-bind_rows(splitpc)
return(chains)}
```

## Alpha

The daily intensity of interactions among hummingbirds and plants

Grab 1 species to visualize

```{r,fig.height=8,fig.width=8}
chains$intercept<-getChains(models$intercept)

#merge species names
chains$intercept<-chains$intercept %>% left_join(jagsIndexBird) %>% left_join(jagsIndexPlants)

#alpha
chains$intercept %>% filter(parameter %in% "alpha",jPlant %in% c(5,10)) %>% ggplot(.,aes(x=exp(value),fill=Plant)) + geom_histogram() + facet_wrap(~Hummingbird,scales="free")
```

```{r,fig.height=8,fig.width=8}
chains$intercept %>% filter(parameter %in% "alpha") %>% group_by(Hummingbird,Plant) %>% summarize(n=mean(value)) %>% ggplot(.,aes(fill=inv.logit(n),x=Plant,y=Hummingbird)) + geom_tile() + scale_fill_continuous(low="blue",high="red") + theme(axis.text.x = element_text(angle=-90)) + labs(fill="P(Interaction)")
```

## Lambda

The phylogenetic signal in interaction probability

```{r}
chains$intercept %>% filter(parameter %in% "lambda") %>% ggplot(.,aes(x=Draw,y=value,col=as.factor(chain))) + geom_line() 

chains$intercept %>%  filter(parameter %in% "lambda") %>% ggplot(.,aes(x=value,fill=par)) + geom_histogram()

```

## sigma

The variance phylogenetic signal in interaction probability

```{r}
chains$intercept %>% filter(parameter %in% "sigma") %>% ggplot(.,aes(x=Draw,y=value,col=as.factor(chain))) + geom_line() 

chains$intercept %>%  filter(parameter %in% "sigma") %>% ggplot(.,aes(x=value,fill=par)) + geom_histogram()
```

# e

This can be thought of as the effect size of phylogenetic influence. Change in probability of interactions due to interactions

```{r}
chains$intercept %>%  filter(parameter %in% "e") %>% ggplot(.,aes(fill=Plant,x=inv.logit(value))) + geom_histogram() + labs(fill="Probability of interactions")
```


```{r}
save.image("/Users/ben/Dropbox/Gesner/analysis.RData")
```
