---
title: "Gesneriaceae Visitor Overlap"
author: "Ben Weinstein"
date: "May 8, 2017"
  html_document:
    number_sections: true
    toc: true
---

```{r}
library(knitr)
opts_chunk$set(echo=F,warning = F,message = F)
```

```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(e1071)
library(ggbiplot)
library(anytime)
library(vegan)
```

## Read in data

```{r}
#hummingbird interactions
int<-read.csv("data/HummingbirdInteractions.csv")

int<-int %>% filter(!is.na(ID))

#flower phenology
fl<-read.csv("data/FlowerTransectClean.csv")

#plant morphology
morph<-read.csv("data/Species_SantaLucia.csv")
#TODO rename rows

#elevation ranges
elevH<-read.csv("data/HummingbirdElevation.csv")
elevP<-read.csv("data/PlantElevation.csv")

#Data cleaning 
int$timestamp<-as.POSIXct(paste(int$Time,int$DateP),format="%H:%M:%S %Y-%m-%d")

#one date error
int[int$DateP %in% '2013-07-25',"Month"]<-7

#one duplicate camera error, perhaps two GPS records.
int<-int[!(int$ID %in% "FH1108" & int$Date_F %in% '2014-12-01'),]

#Correct known taxonomic disagreements, atleast compared to traits
int[int$Iplant_Double=="Alloplectus purpureus","Iplant_Double"]<-"Glossoloma purpureum"
int[int$Iplant_Double=="Capanea affinis","Iplant_Double"]<-"Kohleria affinis"
int[int$Iplant_Double=="Columnea cinerea","Iplant_Double"]<-"Columnea mastersonii"
int[int$Iplant_Double=="Alloplectus teuscheri","Iplant_Double"]<-"Drymonia teuscheri"

fl[fl$Iplant_Double=="Alloplectus tetragonoides","Iplant_Double"]<-"Drymonia collegarum"

#use effective corolla where possible.
morph$Corolla<-morph$EffectiveCorolla

morph[is.na(morph$Corolla),"Corolla"]<-morph[is.na(morph$Corolla),"TotalCorolla"]

#First row is empty
morph<-morph[-1,]

morph<-morph %>% select(Iplant_Double=Group.1,TotalCorolla,EffectiveCorolla,Corolla.Width)

#subset phenology data for gesneriaceae, must have atleast 10 records
fl<-fl %>% select(Date=Date_F,Family,Genus,Species,Iplant_Double,Total_Flowers,Height,Elevation=ele,month,Year) %>% filter(Family=="Gesneriaceae")

tokeep<-fl %>% group_by(Iplant_Double) %>% dplyr::summarize(n=n()) %>% filter(n>20) %>% .$Iplant_Double

fl<-fl %>% filter(Iplant_Double %in% tokeep) %>% filter(!Species=="")
fl<-droplevels(fl)

#Combine levels from transect, morph, elev and camera datasets
fl<-fl %>% filter(Iplant_Double %in% unique(int$Iplant_Double))

int<-int %>% filter(Iplant_Double %in% unique(fl$Iplant_Double)) %>% select(ID,Date=DateP,Iplant_Double,Hummingbird,Sex,Pierce,Elevation=ele,Transect_R)

elevP<-elevP %>% filter(Iplant_Double %in% unique(fl$Iplant_Double)) %>% select(Iplant_Double,Low,m,High)

elevH<-elevH %>% filter(Hummingbird %in% unique(int$Hummingbird)) %>% select(Hummingbird,Low,m,High)

morph<-morph %>% filter(Iplant_Double %in% unique(fl$Iplant_Double))
```

#Interaction Matrix

Binary network to begin with

```{r}
int_binary<-int %>% group_by(Hummingbird,Iplant_Double) %>% dplyr::summarize(n=n())
int_binary$Obs<-(int_binary$n>0)*1

ggplot(int_binary,aes(x=Iplant_Double,y=Hummingbird,fill=as.factor(Obs))) + geom_tile() + theme_bw() + theme(axis.text.x=element_text(angle=-90))  + labs(x="Plants",fill="Visits") + scale_fill_manual(values=c("black"))
```

# Plant Overlap in Hummingbird Visitor

```{r}
hp<-reshape2::acast(int_binary,Iplant_Double~Hummingbird,fill=0)

#no variance
df_f <- hp[,apply(hp, 2, var, na.rm=TRUE) != 0]

ggbiplot::ggbiplot(prcomp(df_f),labels=rownames(hp)) + coord_cartesian(xlim=c(-2.5,2.5),ylim=c(-2.5,2.5)) + theme_bw()
ggsave("Biplot.svg")
```

# Plant Overlap in Time

```{r,fig.height=10,fig.width=12}

fl$Month<-month.name[fl$month]
fl$Month<-factor(fl$Month,levels=month.name)

sumf<-fl %>% group_by(Iplant_Double,Month,Year) %>% dplyr::summarize(n=n())

#TODO fix this ugly function
#expand empty factor to have a 0 count
zero_count<-expand.grid(Iplant_Double=unique(sumf$Iplant_Double),Month=unique(sumf$Month),Year=unique(sumf$Year),n=0)

zero_count<-zero_count %>% anti_join(sumf,by=c("Iplant_Double","Month","Year"))
sumf<-bind_rows(sumf,zero_count)
ggplot(sumf,aes(x=Month,y=n)) + geom_point(size=2)  + geom_smooth(aes(group=1),span = 0.5,se=F) + facet_wrap(~Iplant_Double,scales = "free_y",ncol=4) + theme(axis.text.x=element_text(angle=-90)) + labs("Flowers") + theme_bw() + ylim(0,NA)
ggsave("Phenology.svg",height=6,width=10)
```

#Describing the phenological strategies

We need measures of plant phenology stragies.

* Disperson: The variance in flowers per month
* Consistancy: The variance in flowers per month per year
* Kurtosis: The peakiness of the per year flowering pulse.

To compare species with such difference flowering forms and life history strategies, I think its product to standardize by the total flowers per species

```{r}
sumfz<-sumf %>% group_by(Iplant_Double) %>% mutate(n=n/sum(n))

firstmeasure<-sumfz %>% group_by(Iplant_Double,Year) %>% summarize(Dispersion=var(n,na.rm=T),Kurtosis=kurtosis(n,na.rm=T))

allmeasures<-sumfz %>% group_by(Iplant_Double,Month) %>% summarize(Variability=var(n,na.rm=T)) %>% summarize(Variability=mean(Variability,na.rm=T)) %>% merge(firstmeasure)
ggplot(allmeasures) + geom_point(aes(x=Dispersion,y=Kurtosis,col=Iplant_Double)) + facet_wrap(~Year,scales = "free")

firstmeasure_mean<-sumfz %>% group_by(Iplant_Double,Year) %>% summarize(Dispersion=var(n),Kurtosis=kurtosis(n,na.rm=T)) %>% summarise(Dispersion=mean(Dispersion),Kurtosis=mean(Kurtosis,na.rm=T))
allmeasures_mean<-sumfz %>% group_by(Iplant_Double,Month) %>% summarize(Variability=var(n)) %>% summarize(Variability=mean(Variability,na.rm=T)) %>% merge(firstmeasure_mean)

rownames(allmeasures_mean)<-allmeasures_mean[,1]
allmeasures_mean<-allmeasures_mean[,-1]
ggbiplot(prcomp(allmeasures_mean,scale=T,center=T),labels=rownames(allmeasures_mean)) + theme_bw()
```

# Is the observed pattern non-random with respect to time?

Only the total number of observations are mantained.

We are looking for consistancy in patterns across years.

```{r}

null1<-function(){
  sumf<-fl %>% mutate(Month=sample(month.name,size=nrow(fl),replace=T)) %>% group_by(Iplant_Double,Month,Year) %>% dplyr::summarize(n=n())
  return(sumf)
}
out<-list()

for(x in 1:1000){
  out[[x]]<-data.frame(Iteration=x,null1())
}
null1_result<-bind_rows(out)

null1_sum<-null1_result %>% group_by(Month,Year,Iplant_Double) %>% summarize(lower=quantile(n,0.05),upper=quantile(n,0.95)) 

ggplot(null1_sum) + geom_ribbon(aes(ymin=lower,ymax=upper,x=Month,group=Year)) + facet_wrap(~Iplant_Double,scales="free_y") + geom_point(data=sumf,aes(x=Month,y=n,col=as.factor(Year))) + theme_bw() + labs(col="Year")

#Plot as differences.
null1_diff<-null1_sum %>% merge(sumf) %>% mutate(lower=n-lower,upper=n-upper)
ggplot(null1_diff) + geom_ribbon(aes(ymin=lower,ymax=upper,x=Month,group=Year),alpha=0.4) + facet_wrap(~Iplant_Double,scales="free_y",ncol=3) + theme_bw() + labs(col="Year") + geom_hline(yintercept=0,linetype="dashed",col='red',size=1.1) + scale_x_discrete(labels=month.abb)
ggsave("Figures/Null1.svg")
ggsave("Figures/Null1.jpeg",height=6,width=10)
```

Blooming pattern is non-random with respect to time.

# Is the observed pattern non-random with respect to time and plant abundance.

```{r,eval=F}
null2<-function(){
  sumf<-fl %>% group_by(Iplant_Double) %>% mutate(Month=sample(month.name,size=nrow(.),replace=T)) %>% group_by(Iplant_Double,Month,Year) %>% dplyr::summarize(n=n())
  return(sumf)
}
out<-list()

for(x in 1:1000){
  out[[x]]<-data.frame(Iteration=x,null2())
}
null2_result<-bind_rows(out)

null2_sum<-null2_result %>% group_by(Month,Year,Iplant_Double) %>% summarize(lower=quantile(n,0.05),upper=quantile(n,0.95)) 

ggplot(null2_sum) + geom_ribbon(aes(ymin=lower,ymax=upper,x=Month,group=Year)) + facet_wrap(~Iplant_Double,scales="free_y") + geom_point(data=sumf,aes(x=Month,y=n,col=as.factor(Year))) + theme_bw() + labs(col="Year")

#Plot as differences.
null2_diff<-null2_sum %>% merge(sumf) %>% mutate(lower=n-lower,upper=n-upper)
ggplot(null2_diff) + geom_ribbon(aes(ymin=lower,ymax=upper,x=Month,group=Year),alpha=0.4) + facet_wrap(~Iplant_Double,scales="free_y",ncol=3) + theme_bw() + labs(col="Year") + geom_hline(yintercept=0,linetype="dashed",col='red',size=1.1) + scale_x_discrete(labels=month.abb)
ggsave("Figures/null2.svg")
ggsave("Figures/null2.jpeg",height=6,width=10)
```
