---
title: "Co-flowering and hummingbird overlap in Neotropical Gesneriaceae: Predicting Phenology"
author: "Ben Weinstein"
date: "6/4/2018"
output: 
  html_document:
    toc: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r,message=F,warning=F,include=F}
library(knitr)
opts_chunk$set(message=F,warning=F,echo=F)
library(lubridate)
library(ape)
library(tidyr)
library(kableExtra)
library(boot)
library(dplyr)
library(ggplot2)
library(phytools)
library(stringr)
library(splines)
library(R2jags)
library(truncnorm)

#optionally load
#load("/Users/ben/Dropbox/Gesner/Phenology.RData")

#holder for models and chains
models<-list()
chains<-list()
discrepancy<-list()
```
# Transects
```{r}
#transects
transects<-read.csv("data/cleaned/transects.csv",row.names=1)
head(transects) %>% kable() %>% kable_styling()

#drop NA elevations
transects<-transects %>% filter(!is.na(ele))
```

# Phylogeny

```{r}
phy<-read.newick("data/SantaLucia_tree_Newick")
species<-str_match(phy$tip.label,"_(\\w+)")[,2]
tspecies<-unique(transects$Plant)
to_join<-data.frame(Plant=tspecies,phylo=str_match(tspecies," (\\w+)")[,2])
new_names<-data.frame(phylo=species) %>% inner_join(to_join)
phy$tip.label<- as.character(new_names$Plant)
plot(phy)

#variance covariance matrix - attraction
vcov<-vcv(phy,corr = T)

#variance covariance matrix - attraction
ivcov<-solve(vcov)

#Phylogenetic distance matrix
D<-as.matrix(cophenetic(phy))
D<-round(D/max(D),5)

#Identity matrix
I<-diag(nrow(vcov))

#match transects with phylogeny
transects<-transects %>% filter(Plant %in% phy$tip.label) %>% droplevels()
```

# Traits
```{r}
traits<-read.csv("data/Species_SantaLucia.csv")

#improve column names
traits<-traits %>% select(Plant=Iplant_Double,min_corolla_tube=Lt,corolla_length=Lc,vertical_diameter=D,stamen_length=S,stamen_exertion=E)

#match with phylogeny
traits<-traits %>% filter(Plant %in% phy$tip.label)

#order the factors
head(transects) %>% kable() %>% kable_styling()
```

### Total Flowers

```{r,fig.height=20,eval=F}
transects %>% group_by(Plant,Year) %>%
ggplot(.,aes(x=julian,y=Flowers,col=as.factor(Year))) + geom_point(size=0.1) + geom_line(aes(group=Year)) + facet_wrap(~Plant,scales="free",ncol=2) + labs(col="Year") + theme_bw()
```

## Peak date

```{r}
transects %>% group_by(Plant,Year) %>% filter(Flowers==max(Flowers)) %>% summarize(julian=mean(julian)) %>% arrange(Plant,Year) %>% ggplot(.,aes(x=Plant,y=julian,col=as.factor(Year))) + geom_point(size=3) + theme_bw() + coord_flip() + labs(x="Plant",y="Julian Day of Peak Flowering",col="Year") 
```

```{r,fig.height=10}
ggplot(data=transects) + geom_density(aes(x=julian,fill=Plant)) + facet_wrap(~Plant,ncol=3) + theme_bw()
```

## Infer absences

```{r,fig.height=15}
sumf<-transects %>% group_by(Plant,Date,julian,Year) %>% dplyr::summarize(n=n(),ele=mean(ele)) %>% mutate(Year=as.factor(Year))
zero_count<-expand.grid(Plant=unique(sumf$Plant),Date=unique(sumf$Date),n=0)

#merge elevation for 0 count, get's mean elevation of that day
mean_ele<-transects %>% group_by(Date) %>% summarize(ele=mean(ele))
zero_count<-zero_count %>% anti_join(sumf,by=c("Plant","Date")) %>% mutate(Year=chron::years(as.POSIXct(Date)),julian=yday(as.POSIXct(Date))) %>% inner_join(mean_ele)
transects_zero<-bind_rows(sumf,zero_count)

#transects_zero %>% group_by(Plant,Year) %>%
#ggplot(.,aes(x=as.numeric(julian),y=n,col=as.factor(Year))) + geom_point(size=0.1) + #geom_line(aes(group=as.factor(Year))) + facet_wrap(~Plant,scales="free",ncol=3) + labs(col="Year") + theme_bw()

transects_zero %>% group_by(Plant,Year)  %>% filter(!is.na(n)) %>%
ggplot(.,aes(x=julian,y=n,col=as.factor(Year))) + geom_point(size=2)+  facet_wrap(~Plant,scales="free",ncol=3) + labs(col="Year") + theme_bw()
```

## Species elevation ranges

```{r}
ggplot(transects_zero %>% filter(!n==0),aes(x=Plant,y=ele)) + geom_boxplot() + coord_flip()
```

## Species by transect matrix

```{r}
#first observation of 2017
first_date<-transects_zero %>% ungroup() %>% filter(Year==2017) %>% filter(julian==min(julian)) 

ggplot(transects_zero,aes(x=Plant,y=Date,fill=n>0,col=Year)) + geom_tile()  + 
  coord_flip() + theme(axis.text.x =element_blank()) + labs(y="Observations",x="Species",fill="Flowering")  + scale_fill_manual(values=c("white","black")) 
ggsave("Figures/matrix.svg",height=3,width=11)
```

Check date integrity
```{r}
a<-transects_zero %>% group_by(Date) %>% summarize(min=min(ele),max=max(ele),n=sum(n))
ggplot(a,aes(x=Date)) + geom_linerange(aes(ymin=min,ymax=max))

ggplot(a,aes(x=Date)) + geom_point(aes(y=n,size=n,col=n<5))

a %>% filter(n < 5)

transects_zero %>% filter(Date %in% "2014-01-09")
```

Need to cut into better slices.

# Count model of species phenology

```{r,fig.height=15}
#prepare data
indat<-transects_zero %>% mutate(n=(n>0)*1)

#Easiest to work with jags as numeric ordinal values
indat$Plant<-as.factor(indat$Plant)
indat$jPlant<-as.numeric(indat$Plant)

jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=factor(1:length(levels(indat$Plant))))

#hold 2017 data aside for prediction
to_predict<-indat %>% filter(Year %in% "2017")
indat<-indat %>% filter(!Year %in% "2017")

#give index positions
to_predict$Index<-1:nrow(to_predict)
indat$Index<-1:nrow(indat)
```

```{r}
#Source model
source("model/occ_baseline.R")

#print model
writeLines(readLines("model/occ_baseline.R"))

#Run Model
runModel<-function(Yobs_dat,Ynew_dat){
  
  #Parameters to track
  ParsStage <- c("alpha","fit","fitnew","discrepancy","fitpred","Ynew","prediction")
  
  #Jags Data

  Dat<-list(
    
    #Observed data
    Y=Yobs_dat$n,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    Nobs=length(Yobs_dat$n),

    #Prediction data
    Ypred=Ynew_dat$n,
    Ypred_plant=Ynew_dat$jPlant,
    Npreds=length(Ynew_dat$n))
  
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/occ_baseline.jags",n.thin=1,n.iter=10000,n.burnin=9800,n.chains=2,DIC=F)
    )
    return(model)
}

models$baseline<-runModel(Yobs_dat=indat,Ynew_dat=to_predict)
```

```{r} 
getChains<-function(mod,indat,to_predict){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("alpha")]<-lapply(
    splitpc[c("alpha")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=sv)
    })
  
  #ragged index - observed
  splitpc[c("discrepancy","Ynew")]<-lapply(
    splitpc[c("discrepancy","Ynew")],function(x){
    #get index
    sv<-data.frame(Index=as.numeric(str_match(x$par,"(\\w+)\\[(\\d+)]")[,3]))
    
    #merge with indat to get species
    ragged<-sv %>% left_join(indat[,c("Index","jPlant")]) %>% mutate(jPlant=as.character(jPlant))
    
    pc<-data.frame(x,ragged)
    return(pc)
    })
  
    #ragged index - prediction
  splitpc[c("prediction")]<-lapply(
    splitpc[c("prediction")],function(x){
    #get index
    sv<-data.frame(Index=as.numeric(str_match(x$par,"(\\w+)\\[(\\d+)]")[,3]))
    
    #merge with indat to get species
    ragged<-sv %>% left_join(to_predict[,c("Index","jPlant")]) %>% mutate(jPlant=as.character(jPlant))
    
    pc<-data.frame(x,ragged)
    return(pc)
    })
  chains<-bind_rows(splitpc)
return(chains)}
```

```{r}
chains$baseline<-getChains(models$baseline,indat,to_predict) %>% left_join(jagsIndexPlants)
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$baseline %>% filter(parameter %in% c("alpha")) %>%                    ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$baseline %>% filter(parameter %in% c("fit","fitnew","fitpred")) %>%                    ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=1) + labs(col="chain")
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$baseline %>% filter(parameter %in% c("alpha","fit","fitnew","fitpred")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=1,scales="free")
```

### Error by elevation for estimated data 

```{r,fig.height=11}

preds<-chains$baseline %>% filter(parameter %in% c("discrepancy")) %>% group_by(Plant,Index) %>% mutate(jPlant=as.numeric(as.character(jPlant))) %>% inner_join(indat) %>% group_by(Plant,ele) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(preds) + geom_ribbon(aes(x=ele,ymin=lower,ymax=upper),alpha=0.5)+ facet_wrap(~Plant) 
```

# Phylogeny

## Attraction

```{r,fig.height=15}
#prepare data
indat<-transects_zero %>% filter(Plant %in% phy$tip.label) %>% mutate(n=(n>0)*1)
indat$Plant<-factor(indat$Plant,levels=phy$tip.label)
indat$jPlant<-as.numeric(indat$Plant)
indat$Factor_Date<-as.factor(indat$Date)
indat$jDate<-as.numeric(indat$Factor_Date)
jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=factor(1:length(levels(indat$Plant))))

#hold 2017 data aside for prediction
to_predict<-indat %>% filter(Year %in% "2017")
indat<-indat %>% filter(!Year %in% "2017")

#give index positions
to_predict$Index<-1:nrow(to_predict)
indat$Index<-1:nrow(indat)
```

```{r}
#Source model
source("model/occ_attraction.R")

#print model
writeLines(readLines("model/occ_attraction.R"))

#Run Model
runModel_attraction<-function(Yobs_dat,Ynew_dat,D=D){
  
  #Parameters to track
  ParsStage <- c("alpha","fit","fitnew","discrepancy","lambda_cov","e","omega","fitpred","Ynew","prediction","gamma")
  
  #Jags Data

  Dat<-list(
    Y=Yobs_dat$n,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    Nobs=length(Yobs_dat$n),
    zeros=rep(0,max(c(Yobs_dat$jPlant))),
    Date=Yobs_dat$jDate,
    Dates=max(c(Ynew_dat$jDate,Yobs_dat$jDate)),
    D=D,
    I=I,
    
    #Prediction data
    Ypred=Ynew_dat$n,
    Ypred_plant=Ynew_dat$jPlant,
    Ypred_date=Ynew_dat$jDate,
    Npreds=length(Ynew_dat$n))
    
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/occ_attraction.jags",n.thin=1,n.iter=10000,n.burnin=9800,n.chains=2,DIC=F)
    )
    return(model)
}

models$phylogenetic_attraction<-runModel_attraction(Yobs_dat=indat,Ynew_dat=to_predict,D=D)
```

#Get Chains
```{r} 
getChains<-function(mod,indat,to_predict){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("e","alpha")]<-lapply(
    splitpc[c("e","alpha")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=sv)
    })
  
  #ragged index - observed
  splitpc[c("discrepancy","Ynew")]<-lapply(
    splitpc[c("discrepancy","Ynew")],function(x){
    #get index
    sv<-data.frame(Index=as.numeric(str_match(x$par,"(\\w+)\\[(\\d+)]")[,3]))
    
    #merge with observed data to get species
    ragged<-sv %>% left_join(indat[,c("Index","jPlant")]) %>% mutate(jPlant=as.character(jPlant))
    
    pc<-data.frame(x,ragged)
    return(pc)
    })
  
  #ragged index - predicted
  splitpc[c("prediction")]<-lapply(
    splitpc[c("prediction")],function(x){
    
      #get index
    sv<-data.frame(Index=as.numeric(str_match(x$par,"(\\w+)\\[(\\d+)]")[,3]))
    
    #merge with predicted data to get species
    ragged<-sv %>% left_join(to_predict[,c("Index","jPlant")]) %>% mutate(jPlant=as.character(jPlant))
    
    pc<-data.frame(x,ragged)
    return(pc)
    })
  chains<-bind_rows(splitpc)
return(chains)}
```

```{r}
chains$phylogenetic_attraction<-getChains(models$phylogenetic_attraction,indat,to_predict) %>%  left_join(jagsIndexPlants)
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$phylogenetic_attraction %>% filter(parameter %in% c("alpha","e")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$phylogenetic_attraction %>% filter(parameter %in% c("fit","fitnew","lambda_cov","omega","fitpred","gamma")) %>%  ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=2)
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$phylogenetic_attraction %>% filter(!parameter %in% c("discrepancy","Ynew","prediction")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

## Decay in phylogenetic attraction

```{r}

effect<-function(lambda,D){
  y<-as.numeric(exp(-lambda*D))
  data.frame(lambda,Distance=as.numeric(D),Covariance=y)
}

phy_effect<-chains$phylogenetic_attraction %>% filter(parameter=="lambda_cov") %>% group_by(Draw,chain) %>% do(effect(.$value,D)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(phy_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Phylogenetic Covariance")
```

## Repulsion

```{r}
#Source model
source("model/occ_repulsion.R")

#print model
writeLines(readLines("model/occ_repulsion.R"))

#Run Model
runModel_repulsion<-function(Yobs_dat,Ynew_dat,D=D){
  
  
  #Parameters to track
  ParsStage <- c("alpha","fit","fitnew","discrepancy","lambda_cov","e","omega","fitpred","Ynew","prediction","gamma")
  
  #Jags Data

  Dat<-list(
    Y=Yobs_dat$n,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    Nobs=length(Yobs_dat$n),
    zeros=rep(0,max(c(Yobs_dat$jPlant))),
    Date=Yobs_dat$jDate,
    Dates=max(c(Ynew_dat$jDate,Yobs_dat$jDate)),
    D=D,
    I=I,
    
    #Prediction data
    Ypred=Ynew_dat$n,
    Ypred_plant=Ynew_dat$jPlant,
    Ypred_date=Ynew_dat$jDate,
    Npreds=length(Ynew_dat$n))
  
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/occ_repulsion.jags",n.thin=1,n.iter=10000,n.burnin=9800,n.chains=2,DIC=F)
    )
    return(model)
}

models$phylogenetic_repulsion<-runModel_repulsion(Yobs_dat=indat,Ynew_dat=to_predict,D=D)
```

#Get Chains

```{r}
chains$phylogenetic_repulsion<-getChains(models$phylogenetic_repulsion,indat,to_predict) %>%  left_join(jagsIndexPlants)
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$phylogenetic_repulsion %>% filter(parameter %in% c("alpha","e")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$phylogenetic_repulsion %>% filter(par %in% c("fit","fitnew","lambda_cov","omega","fitpred","gamma")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=2)
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$phylogenetic_repulsion %>% filter(!parameter %in% c("discrepancy","Ynew","prediction")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

## Decay in phylogenetic repulsion

```{r}

effect_repulsion<-function(lambda,D){
  y<-as.numeric(solve(exp(-lambda*D)))
  data.frame(lambda,Distance=as.numeric(D),Covariance=y)
}

phyrepulse_effect<-chains$phylogenetic_repulsion %>% filter(parameter=="lambda_cov") %>% group_by(Draw,chain) %>% do(effect_repulsion(.$value,D)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(phyrepulse_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```

# Traits

## Trait Attraction

```{r,fig.height=15}
#prepare data
indat<-transects_zero %>% filter(Plant %in% phy$tip.label) %>% mutate(n=(n>0)*1)

```

```{r}
#Distance matrix
rownames(traits)<-traits[,1]
Dtraits<-as.matrix(dist(traits[,-1]))
Dtraits<-Dtraits/max(Dtraits)

#filter data to match traits
indat<-transects_zero %>% filter(Plant %in% rownames(traits)) %>% mutate(n=(n>0)*1)
indat$Plant<-factor(indat$Plant,levels=rownames(traits))
indat$jPlant<-as.numeric(indat$Plant)
indat$Factor_Date<-as.factor(indat$Date)
indat$jDate<-as.numeric(indat$Factor_Date)
jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=factor(1:length(levels(indat$Plant))))

#hold 2017 data aside for prediction
to_predict<-indat %>% filter(Year %in% "2017")
indat<-indat %>% filter(!Year %in% "2017")

#give index positions
to_predict$Index<-1:nrow(to_predict)
indat$Index<-1:nrow(indat)
```

```{r}
#Source model
source("model/occ_attraction.R")

#print model
writeLines(readLines("model/occ_attraction.R"))

models$trait_attraction<-runModel_attraction(Yobs_dat=indat,Ynew_dat=to_predict,D=Dtraits)
```

#Get Chains

```{r}
chains$trait_attraction<-getChains(models$trait_attraction,indat,to_predict) %>%  left_join(jagsIndexPlants)
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$trait_attraction %>% filter(parameter %in% c("alpha","e")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$trait_attraction %>% filter(par %in% c("fit","fitnew","lambda_cov","fitpred","omega","gamma")) %>%  ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=2)
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$trait_attraction %>% filter(!parameter %in% c("discrepancy","Ynew","prediction")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

## Decay in trait attraction

```{r}
trait_effect<-chains$trait_attraction %>% filter(parameter=="lambda_cov") %>% group_by(Draw,chain) %>% do(effect(.$value,Dtraits)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(trait_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```

## Repulsion

```{r}
#Source model
source("model/occ_repulsion.R")

#print model
writeLines(readLines("model/occ_repulsion.R"))

models$trait_repulsion<-runModel_repulsion(Yobs_dat=indat,Ynew_dat=to_predict,D=Dtraits)
```

#Get Chains

```{r}
chains$trait_repulsion<-getChains(models$trait_repulsion,indat,to_predict) %>%  left_join(jagsIndexPlants)
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$trait_repulsion %>% filter(parameter %in% c("alpha","e")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$trait_repulsion %>% filter(par %in% c("fit","fitnew","lambda_cov","omega","fitpred","gamma")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=2)
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$trait_repulsion %>% filter(!parameter %in% c("discrepancy","Ynew","prediction")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

## Decay in trait repulsion

```{r}
traitrepulse_effect<-chains$trait_repulsion %>% filter(parameter=="lambda_cov") %>% group_by(Draw,chain) %>% do(effect_repulsion(.$value,Dtraits)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(traitrepulse_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```

# Model Comparison

```{r}
#reduce models to a single df
for(x in 1:length(chains)){
  chains[[x]]$Model<-names(chains)[[x]]
}

allchains<-bind_rows(chains)

#create an attraction/repulsion label
allchains$Effect<-str_detect(allchains$Model,"repulsion")

allchains[!allchains$Effect==TRUE,"Effect"]<-"Attraction"
allchains[allchains$Effect==TRUE,"Effect"]<-"Repulsion"
allchains[allchains$Model=="julian","Effect"]<-NA
```

## Alpha

The probability of occurrence.

```{r}
#Alphas

#The baseline model is not logit transformed intercept
alphas<-allchains %>% filter(parameter %in% "alpha")

alphas[!alphas$Model %in% "baseline","value"]<-inv.logit(alphas[!alphas$Model %in% "baseline","value"])

alphas %>% group_by(Model,Plant) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% ggplot(.,aes(x=Model,y=mean,col=Plant)) + geom_pointrange(aes(ymin=lower,ymax=upper)) + coord_flip() + facet_wrap(~Plant,scales="free")
```

## E: The effect of autocorrelation on mean flowering intensity

```{r,fig.height=12,fig.width=10}
allpars %>% filter(parameter %in% "e") %>% ggplot(.,aes(x=Model,y=mean)) + geom_pointrange(aes(ymin=lower,ymax=upper)) + facet_wrap(~Plant,scales="free",ncol=3) + coord_flip() + labs(y="E")
```

## Omega: The magnitude of the effect of autocorrelation on mean flowering occurrence

```{r}
allpars %>% filter(parameter %in% "omega") %>% ggplot(.,aes(x=Model,y=mean)) + geom_pointrange(aes(ymin=lower,ymax=upper)) + coord_flip() + labs(y="Omega")
```

## Gamma: The variance of the effect of autocorrelation on mean flowering occurrence

```{r,fig.height=12,fig.width=10}
allpars %>% filter(parameter %in% "gamma") %>% ggplot(.,aes(x=Model,y=mean)) + geom_pointrange(aes(ymin=lower,ymax=upper)) + coord_flip() + labs(y="Gamma")
```

## Effect of autocorrelation

TODO, e is not indexed correctly yet.

```{r}
autocorrelation<-allchains %>% select(-par) %>% filter(parameter %in% "e")  %>% group_by(Model,Plant) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(autocorrelation) + geom_pointrange(aes(x=Plant,ymin=lower,ymax=upper,y=mean,col=Model))+ coord_flip() + labs(y="Effect on intensity")
```

## Decay in autocorrelation effect

```{r}
phy_effect$Model<-"Phylogenetic Attraction"
phyrepulse_effect$Model<-"Phylogenetic Repulsion"
trait_effect$Model<-"Trait Attraction"
traitrepulse_effect$Model<-"Trait Repulsion"

est_effect<-bind_rows(list(phy_effect,phyrepulse_effect,trait_effect,traitrepulse_effect))
ggplot(est_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance") + facet_wrap(~Model,scales="free")
```

# Model Fit

## Bayesian pvalue

```{r,fig.height=10,fig.width=10}

fitstat<-allchains %>% filter(parameter %in% c("fit","fitnew")) %>% select(Model,Effect,Draw,chain,parameter,value)  %>% spread(key=parameter,value=value) 

ymin<-min(fitstat$fit)
ymax<-max(fitstat$fit)
ab<-data.frame(x=seq(0,ymax,1),y=seq(0,ymax,1))

ggplot(data=fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data",col="Model")  + ggtitle("Posterior predictive check") + coord_fixed() + geom_line(data=ab,aes(x=x,y=y),group=1,linetype="dashed") 

#Bayesian p-value
fitstat %>% group_by(Model) %>% summarize(p=sum((fitnew>fit)/n()))
```

## Overall

```{r}
allchains %>% filter(parameter %in% "fit") %>% mutate(value=value) %>%  group_by(Model) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% ggplot(.,aes(y=mean,x=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Discrepancy") + theme_bw()
```

## By Species

```{r}
allchains %>% filter(parameter %in% "discrepancy") %>% group_by(Draw,chain,Model,Plant) %>% summarize(value=sqrt(sum(value)/n())) %>% group_by(Model,Plant) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% arrange(mean) %>% ggplot(.,aes(x=Plant,y=mean,col=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Discrepancy") + theme_bw()
```

Zoom in 
```{r,fig.height=10,fig.width=12}
allchains %>% filter(parameter %in% "discrepancy") %>% group_by(Draw,chain,Model,Plant) %>% summarize(value=sqrt(sum(value)/n())) %>% group_by(Model,Plant) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% arrange(mean) %>% ggplot(.,aes(y=mean,x=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + labs(y="Root Mean Squared Error") + theme_bw() + facet_wrap(~Plant,ncol=3,scales="free") + coord_flip()
```
#Prediction

```{r}
allchains %>% filter(parameter %in% "fitpred") %>% mutate(value=value) %>%  group_by(Model) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% ggplot(.,aes(y=mean,x=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Root Mean Squared Error") + theme_bw()
```

## Out of sample data

```{r,fig.height=11,fig.width=12}
preds<-allchains %>% filter(parameter %in% c("prediction")) %>% group_by(Model,Plant,Index)

ggplot(preds) + geom_density(aes(x=value,fill=Model)) + facet_wrap(~Plant,scales="free",ncol=3) + labs(x="Julian Day",y="Density") + geom_density(data=to_predict,aes(x=n),alpha=0.5,fill="black") + theme_bw()
```

```{r}
save.image("/Users/ben/Dropbox/Gesner/Phenology.RData")
```
