---
title: "Co-flowering and hummingbird overlap in Neotropical Gesneriaceae: Predicting Phenology"
author: "Ben Weinstein"
date: "6/4/2018"
output: 
  html_document:
    toc: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r,message=F,warning=F,include=F}
library(knitr)
opts_chunk$set(message=F,warning=F,echo=F)
library(lubridate)
library(ape)
library(tidyr)
library(kableExtra)
library(boot)
library(dplyr)
library(ggplot2)
library(phytools)
library(stringr)
library(splines)
library(R2jags)

#optionally load
#load("/Users/ben/Dropbox/Gesner/Phenology.RData")

#holder for models and chains
models<-list()
chains<-list()
discrepancy<-list()
```
# Transects
```{r}
#transects
transects<-read.csv("data/cleaned/transects.csv",row.names=1)
head(transects) %>% kable() %>% kable_styling()

```

# Phylogeny

```{r}
phy<-read.newick("data/SantaLucia_tree_Newick")
species<-str_match(phy$tip.label,"_(\\w+)")[,2]
tspecies<-unique(transects$Plant)
to_join<-data.frame(Plant=tspecies,phylo=str_match(tspecies," (\\w+)")[,2])
new_names<-data.frame(phylo=species) %>% inner_join(to_join)
phy$tip.label<- as.character(new_names$Plant)
plot(phy)

#variance covariance matrix - attraction
vcov<-vcv(phy,corr = T)

#variance covariance matrix - attraction
ivcov<-solve(vcov)

#Phylogenetic distance matrix
D<-as.matrix(cophenetic(phy))
D<-round(D/max(D),4)

#Identity matrix
I<-diag(nrow(vcov))
```

# Traits
```{r}
traits<-read.csv("data/Species_SantaLucia.csv")

#improve column names
traits<-traits %>% select(Plant=Iplant_Double,min_corolla_tube=Lt,corolla_length=Lc,vertical_diameter=D,stamen_length=S,stamen_exertion=E)

#order the factors
head(transects) %>% kable() %>% kable_styling()
```

### Total Flowers

```{r,fig.height=20}
transects %>% group_by(Plant,Year) %>%
ggplot(.,aes(x=julian,y=Flowers,col=as.factor(Year))) + geom_point(size=0.1) + geom_line(aes(group=Year)) + facet_wrap(~Plant,scales="free",ncol=2) + labs(col="Year") + theme_bw()
```

Peak date

```{r}
transects %>% group_by(Plant,Year) %>% filter(Flowers==max(Flowers)) %>% summarize(julian=mean(julian)) %>% arrange(Plant,Year) %>% ggplot(.,aes(x=Plant,y=julian,col=as.factor(Year))) + geom_point(size=3) + theme_bw() + coord_flip() + labs(x="Plant",y="Julian Day of Peak Flowering",col="Year") 
```


#Count model of species phenology

## Infer absences

```{r,fig.height=15}
sumf<-transects %>% group_by(Plant,Date,julian,Year) %>% dplyr::summarize(n=n()) %>% mutate(Year=as.factor(Year))
zero_count<-expand.grid(Plant=unique(sumf$Plant),Date=unique(sumf$Date),n=0)
zero_count<-zero_count %>% anti_join(sumf,by=c("Plant","Date")) %>% mutate(Year=chron::years(as.POSIXct(Date)),julian=yday(as.POSIXct(Date)))
transects_zero<-bind_rows(sumf,zero_count)

#transects_zero %>% group_by(Plant,Year) %>%
#ggplot(.,aes(x=as.numeric(julian),y=n,col=as.factor(Year))) + geom_point(size=0.1) + #geom_line(aes(group=as.factor(Year))) + facet_wrap(~Plant,scales="free",ncol=3) + labs(col="Year") + theme_bw()

transects_zero %>% group_by(Plant,Year)  %>% filter(!is.na(n)) %>%
ggplot(.,aes(x=julian,y=n,col=as.factor(Year))) + stat_smooth() + facet_wrap(~Plant,scales="free",ncol=3) + labs(col="Year") + theme_bw()
```

```{r}
#jags does better with scaled values, divide julian by 365, we can always reverse it.
transects_zero$juliand<-transects_zero$julian/365
```

## Julian Day

```{r,fig.height=15}
#prepare data
indat<-transects_zero

ggplot(indat,aes(x=juliand,y=n)) + geom_point() + facet_wrap(~Plant,scales="free",ncol=3)

#Easiest to work with jags as numeric ordinal values
indat$Plant<-as.factor(indat$Plant)
indat$jPlant<-as.numeric(indat$Plant)

jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=factor(1:length(levels(indat$Plant))))
```

```{r}
#Source model
source("model/julian.R")

#print model
writeLines(readLines("model/julian.R"))

#Run Model
runModel<-function(Yobs_dat){
  
  
  #Parameters to track
  ParsStage <- c("alpha","beta","beta2")
  
  #Jags Data


  Dat<-list(
    Yobs=Yobs_dat$n,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    julian=Yobs_dat$juliand,
    Nobs=length(Yobs_dat$n))
  
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/julian.jags",n.thin=1,n.iter=50000,n.burnin=49500,n.chains=2,DIC=F)
    )
    return(model)
}

models$julian<-runModel(Yobs_dat=indat)
```

#Get Chains
```{r}
getChains<-function(mod){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("alpha","beta","beta2")]<-lapply(
    splitpc[c("alpha","beta","beta2")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=sv)
    })
  
    
  chains<-bind_rows(splitpc)
return(chains)}
```

```{r}
chains$julian<-getChains(models$julian) %>% inner_join(jagsIndexPlants)
```

### Evaluate convergence

```{r,fig.height=15,fig.width=20}
ggplot(chains$julian,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

### Posterior estimates
```{r}
ggplot(chains$julian,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=1,scales="free")
```

### Effect of Julian day

```{r}

predict_plant<-function(alpha,beta,beta2,julian){
  result<-exp(alpha + beta * julian + beta2 * julian^2)
  return(data.frame(alpha,beta,beta2,julian,value=result))
}

days_to_predict<-unique(indat$juliand)
preds<-chains$julian %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict)) %>% group_by(Plant,julian) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(preds,aes(x=julian,y=mean))  + geom_ribbon(aes(ymin=lower,ymax=upper)) + facet_wrap(~Plant,ncol=4,scales="free") + labs(y="Flowering Intensity",x="Julian/365") + theme_bw()
```

```{r}
### Compute discrepancy
discrepancy$julian<-chains$julian %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict))  %>% ungroup() %>% rename(juliand=julian) %>% inner_join(indat[,c("Plant","juliand","n")],by=c("Plant","juliand")) %>% mutate(Discrepancy=abs(n-value))
```

# Phylogeny

## Julian + Phylogenetic Attraction

```{r}
#filter data to match phylogeny
indat<-transects_zero %>% filter(Plant %in% phy$tip.label)
indat$Plant<-factor(indat$Plant,levels=phy$tip.label)
indat$jPlant<-as.numeric(indat$Plant)
jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=factor(1:length(levels(indat$Plant))))

```

```{r}
#Source model
source("model/julian_decay.R")

#print model
writeLines(readLines("model/julian_decay.R"))

#Run Model
runModel<-function(Yobs_dat,D){
  
  
  #Parameters to track
  ParsStage <- c("alpha","beta","beta2","lambda")
  
  InitState<- function(){
    list(
    beta = dat$zeros,
    beta2 = dat$zeros
    )
  }
  
  #Jags Data

  Dat<-list(
    Yobs=Yobs_dat$n,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    julian=Yobs_dat$juliand,
    Nobs=length(Yobs_dat$n),
    D=D,
    zeros=rep(0,max(c(Yobs_dat$jPlant))))
  
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/julian_decay.jags",n.thin=1,n.iter=50000,n.burnin=49500,n.chains=2,DIC=F)
    )
    return(model)
}

models$phylogeny_attraction<-runModel(Yobs_dat=indat,D=D)
```

#Get Chains
```{r}
getChains<-function(mod){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("alpha","beta","beta2")]<-lapply(
    splitpc[c("alpha","beta","beta2")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=sv)
    })
  
    
  chains<-bind_rows(splitpc)
return(chains)}
```

```{r}
chains$phylogeny_attraction<-getChains(models$phylogeny_attraction) %>% left_join(jagsIndexPlants)
```

### Evaluate convergence

```{r}
chains$phylogeny_attraction %>% filter(parameter %in% c("lambda")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=1) + theme_bw() + labs(col="Chain")
```

```{r,fig.height=15,fig.width=20}
chains$phylogeny_attraction %>% filter(!parameter %in% c("lambda")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free") + theme_bw() + labs(col="Chain")
```

### Posterior estimates

```{r}
ggplot(chains$phylogeny_attraction,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

### Effect of Julian day

```{r}

predict_plant<-function(alpha,beta,beta2,julian){
  result<-exp(alpha + beta * julian + beta2 * julian^2)
  return(data.frame(alpha,beta,beta2,julian,value=result))
}

days_to_predict<-unique(indat$juliand)
preds<-chains$phylogeny_attraction %>% select(-par) %>% spread(parameter,value) %>% filter(!is.na(Plant)) %>% group_by(Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict)) %>% group_by(Plant,julian) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(preds,aes(x=julian,y=mean))  + geom_ribbon(aes(ymin=lower,ymax=upper)) + facet_wrap(~Plant,ncol=4,scales="free") + labs(y="Flowering Intensity",x="Julian/365") + theme_bw()
```

```{r}
### Compute discrepancy
discrepancy$phylogeny_attraction<-chains$phylogeny_attraction %>% select(-par) %>% filter(!is.na(Plant))  %>% spread(parameter,value) %>% group_by(Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict))  %>% ungroup() %>% rename(juliand=julian) %>% inner_join(indat[,c("Plant","juliand","n")],by=c("Plant","juliand")) %>% mutate(Discrepancy=abs(n-value))
```

## Decay in phylogenetic effect

```{r}

effect<-function(lambda,D){
  y<-as.numeric(exp(-lambda*D))
  data.frame(lambda,Distance=as.numeric(D),Covariance=y)
}

phy_effect<-chains$phylogeny_attraction %>% filter(parameter=="lambda") %>% group_by(Draw,chain) %>% do(effect(.$value,D)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(phy_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```

## Julian + Phylogenetic Repulsion

```{r}
#filter data to match phylogeny
indat<-transects_zero %>% filter(Plant %in% phy$tip.label)
indat$Plant<-factor(indat$Plant,levels=phy$tip.label)
indat$jPlant<-as.numeric(indat$Plant)
jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=factor(1:length(levels(indat$Plant))))

```

```{r}
#Source model
source("model/julian_decay_repulse.R")

#print model
writeLines(readLines("model/julian_decay_repulse.R"))

#Run Model
runModel<-function(Yobs_dat,D){
  
  
  #Parameters to track
  ParsStage <- c("alpha","beta","beta2","lambda")
  
  InitState<- function(){
    list(
    beta = dat$zeros,
    beta2 = dat$zeros
    )
  }
  
  #Jags Data

  Dat<-list(
    Yobs=Yobs_dat$n,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    julian=Yobs_dat$juliand,
    Nobs=length(Yobs_dat$n),
    D=D,
    zeros=rep(0,max(c(Yobs_dat$jPlant))))
  
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/julian_decay_repulse.jags",n.thin=1,n.iter=50000,n.burnin=49500,n.chains=2,DIC=F)
    )
    return(model)
}

models$phylogeny_repulsion<-runModel(Yobs_dat=indat,D=D)
```

#Get Chains
```{r}
getChains<-function(mod){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("alpha","beta","beta2")]<-lapply(
    splitpc[c("alpha","beta","beta2")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=sv)
    })
  
    
  chains<-bind_rows(splitpc)
return(chains)}
```

```{r}
chains$phylogeny_repulsion<-getChains(models$phylogeny_repulsion) %>% left_join(jagsIndexPlants)
```

### Evaluate convergence

```{r}
chains$phylogeny_repulsion %>% filter(parameter %in% c("lambda")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=1) + theme_bw() + labs(col="Chain")
```

```{r,fig.height=15,fig.width=20}
chains$phylogeny_repulsion %>% filter(!parameter %in% c("lambda")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free") + theme_bw() + labs(col="Chain")
```

### Posterior estimates

```{r}
ggplot(chains$phylogeny_repulsion,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

### Effect of Julian day

```{r}

predict_plant<-function(alpha,beta,beta2,julian){
  result<-exp(alpha + beta * julian + beta2 * julian^2)
  return(data.frame(alpha,beta,beta2,julian,value=result))
}

days_to_predict<-unique(indat$juliand)
preds<-chains$phylogeny_repulsion %>% select(-par) %>% spread(parameter,value) %>% filter(!is.na(Plant)) %>% group_by(Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict)) %>% group_by(Plant,julian) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(preds,aes(x=julian,y=mean))  + geom_ribbon(aes(ymin=lower,ymax=upper)) + facet_wrap(~Plant,ncol=4,scales="free") + labs(y="Flowering Intensity") + theme_bw()
```

```{r}
### Compute discrepancy
discrepancy$phylogeny_repulsion<-chains$phylogeny_repulsion %>% select(-par) %>% filter(!is.na(Plant))  %>% spread(parameter,value) %>% group_by(Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict))  %>% ungroup() %>% rename(juliand=julian) %>% inner_join(indat[,c("Plant","juliand","n")],by=c("Plant","juliand")) %>% mutate(Discrepancy=abs(n-value))
```

## Decay in phylogenetic effect

```{r}

effect<-function(lambda,D){
  y<-as.numeric(solve(exp(-lambda*D)))
  data.frame(lambda,Distance=as.numeric(D),Covariance=y)
}

phyrepulse_effect<-chains$phylogeny_repulsion %>% filter(parameter=="lambda") %>% group_by(Draw,chain) %>% do(effect(.$value,D)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(phyrepulse_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```

# Traits

## Julian + Trait Attraction

```{r}

#Distance matrix
rownames(traits)<-traits[,1]
Dtraits<-as.matrix(dist(traits[,-1]))
Dtraits<-Dtraits/max(Dtraits)

#filter data to match traits
indat<-transects_zero %>% filter(Plant %in% rownames(traits))
indat$Plant<-factor(indat$Plant,levels=rownames(traits))
indat$jPlant<-as.numeric(indat$Plant)
jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=factor(1:length(levels(indat$Plant))))

```

```{r}
#Source model
source("model/julian_decay.R")

#print model
writeLines(readLines("model/julian_decay.R"))

#Run Model
runModel<-function(Yobs_dat,D){
  
  
  #Parameters to track
  ParsStage <- c("alpha","beta","beta2","lambda")
  
  InitState<- function(){
    list(
    beta = dat$zeros,
    beta2 = dat$zeros
    )
  }
  
  #Jags Data

  Dat<-list(
    Yobs=Yobs_dat$n,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    julian=Yobs_dat$juliand,
    Nobs=length(Yobs_dat$n),
    D=D,
    zeros=rep(0,max(c(Yobs_dat$jPlant))))
  
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/julian_decay.jags",n.thin=1,n.iter=50000,n.burnin=49500,n.chains=2,DIC=F)
    )
    return(model)
}

models$trait_attraction<-runModel(Yobs_dat=indat,D=Dtraits)
```

#Get Chains
```{r}
getChains<-function(mod){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("alpha","beta","beta2")]<-lapply(
    splitpc[c("alpha","beta","beta2")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=sv)
    })
  
    
  chains<-bind_rows(splitpc)
return(chains)}
```

```{r}
chains$trait_attraction<-getChains(models$trait_attraction) %>% left_join(jagsIndexPlants)
```

### Evaluate convergence

```{r}
chains$trait_attraction %>% filter(parameter %in% c("lambda")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=1) + theme_bw() + labs(col="Chain")
```

```{r,fig.height=15,fig.width=20}
chains$trait_attraction %>% filter(!parameter %in% c("lambda")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free") + theme_bw() + labs(col="Chain")
```

### Posterior estimates

```{r}
ggplot(chains$trait_attraction,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

### Effect of Julian day

```{r}

predict_plant<-function(alpha,beta,beta2,julian){
  result<-exp(alpha + beta * julian + beta2 * julian^2)
  return(data.frame(alpha,beta,beta2,julian,value=result))
}

days_to_predict<-unique(indat$juliand)
preds<-chains$trait_attraction %>% select(-par) %>% spread(parameter,value) %>% filter(!is.na(Plant)) %>% group_by(Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict)) %>% group_by(Plant,julian) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(preds,aes(x=julian,y=mean))  + geom_ribbon(aes(ymin=lower,ymax=upper)) + facet_wrap(~Plant,ncol=4,scales="free") + labs(y="Flowering Intensity",x="Julian/365") + theme_bw()
```

```{r}
### Compute discrepancy
discrepancy$trait_attraction<-chains$trait_attraction %>% select(-par) %>% filter(!is.na(Plant))  %>% spread(parameter,value) %>% group_by(Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict))  %>% ungroup() %>% rename(juliand=julian) %>% inner_join(indat[,c("Plant","juliand","n")],by=c("Plant","juliand")) %>% mutate(Discrepancy=abs(n-value))
```

## Decay in trait effect

```{r}

effect<-function(lambda,D){
  y<-exp(-lambda*as.numeric(D))
  data.frame(lambda,Distance=as.numeric(D),Covariance=y)
}

trait_effect<-chains$trait_attraction %>% filter(parameter=="lambda") %>% group_by(Draw,chain) %>% do(effect(.$value,Dtraits)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(trait_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```

## Julian + Trait Repulsion

```{r}
#Source model
source("model/julian_decay_repulse.R")

#print model
writeLines(readLines("model/julian_decay_repulse.R"))

#Run Model
runModel<-function(Yobs_dat,D){
  
  
  #Parameters to track
  ParsStage <- c("alpha","beta","beta2","lambda")
  
  InitState<- function(){
    list(
    beta = dat$zeros,
    beta2 = dat$zeros
    )
  }
  
  #Jags Data

  Dat<-list(
    Yobs=Yobs_dat$n,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    julian=Yobs_dat$juliand,
    Nobs=length(Yobs_dat$n),
    D=D,
    zeros=rep(0,max(c(Yobs_dat$jPlant))))
  
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/julian_decay_repulse.jags",n.thin=1,n.iter=50000,n.burnin=49500,n.chains=2,DIC=F)
    )
    return(model)
}

models$trait_repulsion<-runModel(Yobs_dat=indat,D=Dtraits)
```

#Get Chains
```{r}
getChains<-function(mod){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("alpha","beta","beta2")]<-lapply(
    splitpc[c("alpha","beta","beta2")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=sv)
    })
  
    
  chains<-bind_rows(splitpc)
return(chains)}
```

```{r}
chains$trait_repulsion<-getChains(models$trait_repulsion) %>% left_join(jagsIndexPlants)
```

### Evaluate convergence

```{r}
chains$trait_repulsion %>% filter(parameter %in% c("lambda")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=1) + theme_bw() + labs(col="Chain")
```

```{r,fig.height=15,fig.width=20}
chains$trait_repulsion %>% filter(!parameter %in% c("lambda")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free") + theme_bw() + labs(col="Chain")
```

### Posterior estimates

```{r}
ggplot(chains$trait_repulsion,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

### Effect of Julian day

```{r}

predict_plant<-function(alpha,beta,beta2,julian){
  result<-exp(alpha + beta * julian + beta2 * julian^2)
  return(data.frame(alpha,beta,beta2,julian,value=result))
}

days_to_predict<-unique(indat$juliand)
preds<-chains$trait_repulsion %>% select(-par) %>% spread(parameter,value) %>% filter(!is.na(Plant)) %>% group_by(Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict)) %>% group_by(Plant,julian) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(preds,aes(x=julian,y=mean))  + geom_ribbon(aes(ymin=lower,ymax=upper)) + facet_wrap(~Plant,ncol=4,scales="free") + labs(y="Flowering Intensity",x="Julian/365") + theme_bw()
```

## Decay in trait effect

```{r}

effect<-function(lambda,D){
  y<-as.numeric(solve(exp(-lambda*D)))
  data.frame(lambda,Distance=as.numeric(D),Covariance=y)
}

traitrepulse_effect<-chains$trait_repulsion %>% filter(parameter=="lambda") %>% group_by(Draw,chain) %>% do(effect(.$value,Dtraits)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(traitrepulse_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```

```{r}
### Compute discrepancy
discrepancy$trait_repulsion<-chains$trait_repulsion %>% select(-par) %>% filter(!is.na(Plant))  %>% spread(parameter,value) %>% group_by(Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict))  %>% ungroup() %>% rename(juliand=julian) %>% inner_join(indat[,c("Plant","juliand","n")],by=c("Plant","juliand")) %>% mutate(Discrepancy=abs(n-value))
```

# Model Comparison

```{r}
#reduce models to a single df
for(x in 1:length(chains)){
  chains[[x]]$Model<-names(chains)[[x]]
}

allchains<-bind_rows(chains)

#create an attraction/repulsion label
allchains$Effect<-str_detect(allchains$Model,"repulsion")

allchains[allchains$Effect==TRUE,"Effect"]<-"Repulsion"
allchains[!allchains$Effect==TRUE,"Effect"]<-"Attraction"

```

## Effect of julian day

```{r,fig.height=12,fig.width=12}
preds<- allchains %>% select(-par) %>% spread(parameter,value) %>% filter(!is.na(Plant)) %>% group_by(Model,Draw,chain,Plant) %>% do(predict_plant(.$alpha,.$beta,.$beta2,julian=days_to_predict)) %>% group_by(Model,Plant,julian) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(preds,aes(x=julian,y=mean,fill=Model))  + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + facet_wrap(~Plant,ncol=3,scales="free") + labs(y="Flowering Intensity",x="Julian/365") + theme_bw()
```

## Decay in phylogenetic effect

```{r}

phy_effect$Model<-"Phylogenetic Attraction"
phyrepulse_effect$Model<-"Phylogenetic Repulsion"
trait_effect$Model<-"Trait Attraction"
traitrepulse_effect$Model<-"Trait Repulsion"

est_effect<-bind_rows(list(phy_effect,phyrepulse_effect,trait_effect,traitrepulse_effect))
ggplot(est_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance") + facet_wrap(~Model,scales="free")
```

# Model Fit

```{r}
#reduce models to a single df
for(x in 1:length(discrepancy)){
  discrepancy[[x]]$Model<-names(discrepancy)[[x]]
}

discrepancy<-bind_rows(discrepancy)

#create an attraction/repulsion label
discrepancy$effect<-str_detect(discrepancy$Model,"repulsion")
discrepancy[!discrepancy$effect==TRUE,"effect"]<-"Attraction"
discrepancy[discrepancy$effect==TRUE,"effect"]<-"Repulsion"

#Julian gets no label
discrepancy[discrepancy$Model=="julian","effect"]<-""

#label variable
v<-data.frame(Model=c("julian","phylogeny_attraction","phylogeny_repulsion","trait_attraction","trait_repulsion"),variable=c("Julian","Phylogeny","Phylogeny","Trait","Trait"))

discrepancy<-discrepancy %>% inner_join(v) 
```

## Overall

```{r}
discrepancy %>% group_by(variable,effect) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% ggplot(.,aes(x=variable,y=mean,col=effect)) + geom_pointrange(aes(ymin=lower,ymax=upper),position = position_dodge(width=0.2)) + coord_flip() + labs(y="Discrepancy") + theme_bw()
```

## By Species

```{r}
discrepancy%>% group_by(Model,Plant) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% ggplot(.,aes(x=Plant,y=mean,col=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Discrepancy") + theme_bw()
```


#Prediction

Todo

```{r}
save.image("/Users/ben/Dropbox/Gesner/Phenology.RData")
```
