---
title: "Co-flowering and hummingbird overlap in Neotropical Gesneriaceae: Predicting Phenology"
author: "Ben Weinstein"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r,message=F,warning=F,include=F}
library(knitr)
opts_chunk$set(message=F,warning=F,echo=F)
library(lubridate)
library(ape)
library(kableExtra)
library(boot)
library(ggplot2)
library(phytools)
library(stringr)
library(ggbiplot)
library(splines)
library(R2jags)
library(truncnorm)
library(tidyr)
library(dplyr)

#optionally load
#load("/Users/ben/Dropbox/Gesner/Phenology.RData")

#holder for models and chains
models<-list()
chains<-list()
discrepancy<-list()
```

# Transects
```{r}
#transects
transects<-read.csv("data/cleaned/transects.csv",row.names=1)
head(transects) %>% kable() %>% kable_styling()

#drop NA elevations
transects<-transects %>% filter(!is.na(ele))
```

# Interactions

```{r}
intdata<-as.matrix(read.csv("PInteraction.csv",row.names = 1))
Dint<-as.matrix(dist(t(intdata)))

rownames(Dint)<-str_replace(rownames(Dint),"\\."," ")
colnames(Dint)<-str_replace(colnames(Dint),"\\."," ")

ggbiplot(prcomp(intdata))
Dint<-Dint/max(Dint)
```

# Phylogeny

```{r}
phy<-read.newick("data/SantaLucia_tree_Newick")
species<-str_match(phy$tip.label,"_(\\w+)")[,2]
tspecies<-unique(transects$Plant)
to_join<-data.frame(Plant=tspecies,phylo=str_match(tspecies," (\\w+)")[,2])
new_names<-data.frame(phylo=species) %>% inner_join(to_join)
phy$tip.label<- as.character(new_names$Plant)
plot(phy)

#variance covariance matrix - attraction
vcov<-vcv(phy,corr = T)

#variance covariance matrix - attraction
ivcov<-solve(vcov)

#Phylogenetic distance matrix
D<-as.matrix(cophenetic(phy))
D<-round(D/max(D),5)

#Identity matrix
I<-diag(nrow(vcov))

#match transects with phylogeny
transects<-transects %>% filter(Plant %in% phy$tip.label) %>% droplevels()
```

# Traits
```{r}
traits<-read.csv("data/Species_SantaLucia.csv")

#improve column names
traits<-traits %>% select(Plant=Iplant_Double,min_corolla_tube=Lt,corolla_length=Lc,vertical_diameter=D,stamen_length=S,stamen_exertion=E)

#match with phylogeny
traits<-traits %>% filter(Plant %in% phy$tip.label)

#order the factors
head(transects) %>% kable() %>% kable_styling()
```

### Total Flowers

```{r,fig.height=20,eval=F}
transects %>% group_by(Plant,Year) %>%
ggplot(.,aes(x=julian,y=Flowers,col=as.factor(Year))) + geom_point(size=0.1) + geom_line(aes(group=Year)) + facet_wrap(~Plant,scales="free",ncol=2) + labs(col="Year") + theme_bw()
```

## Peak date

```{r}
transects %>% group_by(Plant,Year) %>% filter(Flowers==max(Flowers)) %>% dplyr::summarize(julian=mean(julian)) %>% arrange(Plant,Year) %>% ggplot(.,aes(x=Plant,y=julian,col=as.factor(Year))) + geom_point(size=3) + theme_bw() + coord_flip() + labs(x="Plant",y="Julian Day of Peak Flowering",col="Year") 
```

```{r,fig.height=10}
ggplot(data=transects) + geom_density(aes(x=julian,fill=Plant)) + facet_wrap(~Plant,ncol=3) + theme_bw()
```

As range

```{r,fig.height=10}

transects %>% group_by(Plant,Year) %>% summarize(mean=mean(julian),lower=quantile(julian,0.05),upper=quantile(julian,0.95)) %>% ggplot(.,aes(x=Plant,col=as.factor(Year))) + geom_pointrange(aes(y=mean,min=lower,max=upper),size=0.3,position = position_jitter(height = 0.1)) + coord_flip() + labs(y="Julian Day",col="Year") + theme_bw()
```

```{r}
transects %>% group_by(Plant,julian) %>% summarise(n=sum(Flowers)) %>% filter(!is.na(n)) %>%
ggplot(.,aes(x=julian,y=n)) + geom_point(size=2) + labs(col="Year") + theme_bw() + facet_wrap(~Plant,scales = "free_y")
```

## Infer absences

```{r,fig.height=15}

transects$Year<-as.character(format(as.POSIXlt(transects$Date),"%Y"))
transects$Month<-as.numeric(format(as.POSIXlt(transects$Date),"%m"))
transects$Site<-as.numeric(as.factor(cut(transects$ele,seq(1300,2500,200))))

sumf<-transects %>% group_by(Plant,Site,Month,Year) %>% dplyr::summarize(n=n(),ele=mean(ele)) %>% mutate(Year=as.factor(Year))

all_plants<-function(dat){
  data.frame(dat,Plant=unique(sumf$Plant),n=0)
}

#Which combinations are sampled.
sampled<-sumf %>% ungroup() %>% distinct(Site,Month,Year) %>% arrange(Site,Month,Year) %>% group_by(Site,Month,Year) %>% do(all_plants(.))

#merge elevation for 0 count, get's mean elevation of that day
mean_ele<-transects %>% group_by(Site,Month,Year) %>% summarize(ele=mean(ele))
zero_count<-sampled %>% anti_join(sumf,by=c("Plant","Site","Month","Year")) %>% inner_join(mean_ele)


transects_zero<-bind_rows(sumf,zero_count)


```

## Species elevation ranges

```{r}
ggplot(transects_zero %>% filter(!n==0),aes(x=Plant,y=ele)) + geom_boxplot() + coord_flip()
```

## Data Matrix

```{r}

transects_zero<-transects_zero %>% ungroup() %>% mutate(Index=group_indices(.,Site,Month,Year))
ggplot(transects_zero) + geom_tile(aes(x=Index,y=Plant,fill=as.factor(n>0))) + labs(fill="Flowering")  + theme_bw()
```

## Species by transect matrix

# Count model of species phenology

```{r,fig.height=15}
#prepare data
indat<-transects_zero %>% mutate(n=(n>0)*1)

#Easiest to work with jags as numeric ordinal values
indat$Plant<-as.factor(indat$Plant)
indat$jPlant<-as.numeric(indat$Plant)

jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=1:length(levels(indat$Plant)))

#smmoth data
indat<-indat %>% group_by(Plant,jPlant,Site,Month,Year) %>% summarize(n=max(n))

#hold 2017 data aside for prediction
to_predict<-indat %>% filter(Year %in% "2017")
indat<-indat %>% filter(!Year %in% "2017")

#Index for jags
indat$Index<-as.factor(1:nrow(indat))
to_predict$Index<-as.factor(1:nrow(to_predict))
```


```{r}
#Source model
source("model/threshold_baseline.R")

#print model
writeLines(readLines("model/threshold_baseline.R"))

#Run Model
runModel<-function(Yobs_dat,Ynew_dat){
  
  #Parameters to track
  ParsStage <- c("alpha","fit","fitnew","discrepancy","pred_error","fitpred","Ynew")
  
  #Jags Data


  Dat<-list(
    Nobs=length(Yobs_dat$n),
    Y=Yobs_dat$n,
    Plants=max(Yobs_dat$jPlant),
    Plant=Yobs_dat$jPlant,

    #Prediction data
    Npreds=length(Ynew_dat$n),
    Ypred=Ynew_dat$n,
    PredPlant=Ynew_dat$jPlant)

  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/threshold_baseline.jags",n.thin=2,n.iter=20000,n.burnin=19500,n.chains=2,DIC=F)
    )
    return(model)
}

models$baseline<-runModel(Yobs_dat=indat,Ynew_dat=to_predict)
```

#Get Chains
```{r} 
getChains<-function(mod,indat,to_predict){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("alpha")]<-lapply(
    splitpc[c("alpha")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=as.numeric(as.character(sv)))  %>% left_join(jagsIndexPlants)
    })
  
    #double index - observed
  splitpc[c("Ynew","discrepancy")]<-lapply(
    splitpc[c("Ynew","discrepancy")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Index=sv)
    pc<-pc %>% inner_join(indat)
    })
  
  #Prediction
    splitpc[c("pred_error")]<-lapply(
    splitpc[c("pred_error")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Index=sv)
    pc<-pc %>% inner_join(to_predict)
    })
  
  chains<-bind_rows(splitpc)
return(chains)}
```

```{r}
chains$baseline<-getChains(models$baseline,indat,to_predict)
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$baseline %>% filter(parameter %in% c("alpha")) %>%                    ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$baseline %>% filter(parameter %in% c("fit","fitnew","fitpred")) %>%                    ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=1) + labs(col="chain")
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$baseline %>% filter(parameter %in% c("alpha","fit","fitnew","fitpred")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=1,scales="free")
```

# Phylogeny

## Attraction

```{r,fig.height=15}
#prepare data
indat<-transects_zero %>% filter(Plant %in% phy$tip.label) %>%  mutate(n=(n>0)*1)
indat$Plant<-factor(indat$Plant,levels=phy$tip.label)
indat$jPlant<-as.numeric(indat$Plant)
jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=1:length(levels(indat$Plant)))

#hold 2017 data aside for prediction
to_predict<-indat %>% filter(Year %in% "2017")
indat<-indat %>% filter(!Year %in% "2017")

#Add index
indat$Index<-as.factor(1:nrow(indat))
to_predict$Index<-as.factor(1:nrow(to_predict))
```

```{r}
#Run Model
runModel_autocorrelation<-function(Yobs_dat,Ynew_dat,D=D,mod_name){
  
#Source model
source(paste("model/",mod_name,".R",sep=""))

#print model
writeLines(readLines(paste("model/",mod_name,".R",sep="")))

  #Parameters to track
  ParsStage <- c("alpha","fit","fitnew","discrepancy","pred_error","lambda_cov","e","omega","fitpred","Ynew","gamma")
   
  #Jags Data

    Dat<-list(
    Nobs=length(Yobs_dat$n),
    Y=Yobs_dat$n,
    Plants=max(Yobs_dat$jPlant),
    Plant=Yobs_dat$jPlant,
    Months=max(Yobs_dat$Month),
    Month=Yobs_dat$Month,
    Sites=max(Yobs_dat$Site),
    Site=Yobs_dat$Site,
    
    #Autocorrelation data
    zeros=rep(0,max(Yobs_dat$jPlant)),
    D=D,
    I=I,
    
    #Prediction data
    Npreds=length(Ynew_dat$n),
    Ypred=Ynew_dat$n,
    NewPlant=Ynew_dat$jPlant,
    NewSite=Ynew_dat$Site,
    NewMonth=Ynew_dat$Month)
    
    
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file=paste("model/",mod_name,".jags",sep=""),n.thin=2,n.iter=20000,n.burnin=19500,n.chains=2,DIC=F)
    )
    return(model)
}

models$phylogenetic_attraction<-runModel_autocorrelation(Yobs_dat=indat,Ynew_dat=to_predict,D=D,mod_name = "threshold_attraction")
```

#Get Chains
```{r} 
getChains<-function(mod,indat,to_predict){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("alpha")]<-lapply(
    splitpc[c("alpha")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=as.numeric(as.character(sv)))  %>% left_join(jagsIndexPlants)
    })
  
      #double index - observed
  splitpc[c("Ynew","discrepancy")]<-lapply(
    splitpc[c("Ynew","discrepancy")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Index=sv)
    pc<-pc %>% inner_join(indat)
    })
  
  #Prediction
    splitpc[c("pred_error")]<-lapply(
    splitpc[c("pred_error")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Index=sv)
    pc<-pc %>% inner_join(to_predict)
    })
    
    #triple index
  splitpc[c("e")]<-lapply(
    splitpc[c("e")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+)]"))[,3:5]
    colnames(sv)<-c("jPlant","Site","Month")
    sv$jPlant<-as.numeric(as.character(sv$jPlant))
    sv$Site<-as.numeric(as.character(sv$Site))
    sv$Month<-as.numeric(as.character(sv$Month))

    pc<-data.frame(x,sv)
    })
  
  chains<-bind_rows(splitpc)
return(chains)}
```

```{r}
chains$phylogenetic_attraction<-getChains(models$phylogenetic_attraction,indat,to_predict)
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$phylogenetic_attraction %>% filter(parameter %in% c("alpha")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$phylogenetic_attraction %>% filter(parameter %in% c("fit","fitnew","lambda_cov","omega","fitpred","gamma")) %>%  ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=2)
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$phylogenetic_attraction %>% filter(!parameter %in% c("discrepancy","Ynew","pred_error")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

Mean phylogenetic covariance

```{r}
tauC<-function(omega,gamma,I,D,lambda_cov=5){
      C = exp(-lambda_cov * D)
      vCov=omega*C + (1-omega) * I
      result<-vCov*gamma
      result<-reshape2::melt(result)
      return(result)
}

result<-chains$phylogenetic_attraction %>% filter(parameter %in% c("omega","gamma")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(tauC(omega=.$omega,gamma=.$gamma,I=I,D=D)) %>% group_by(Var1,Var2) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(result,aes(x=Var1,y=Var2,fill=mean)) + geom_tile() + scale_fill_continuous(low="white",high="red")
```

## Decay in phylogenetic attraction

```{r}
effect<-function(lambda,D,omega,gamma){
  C<-exp(-lambda*D)
  vCov=(omega*C[,] + (1-omega) * I)
  vCov=vCov*gamma
  data.frame(lambda,Distance=as.numeric(D),Covariance=as.numeric(vCov))
}

phy_effect<-chains$phylogenetic_attraction %>% filter(parameter %in% c("lambda_cov","omega","gamma")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(effect(lambda=.$lambda_cov,D=D,omega=.$omega,gamma=.$gamma)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(phy_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")

```

## Repulsion

```{r}
models$phylogenetic_repulsion<-runModel_autocorrelation(Yobs_dat=indat,Ynew_dat=to_predict,D=D,mod_name = "threshold_repulsion")
```

#Get Chains

```{r}
chains$phylogenetic_repulsion<-getChains(models$phylogenetic_repulsion,indat,to_predict)
```

### Evaluate convergence

```{r,fig.height=10,fig.width=12}
chains$phylogenetic_repulsion %>% filter(parameter %in% c("alpha")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$phylogenetic_repulsion %>% filter(par %in% c("fit","fitnew","lambda_cov","omega","fitpred","gamma")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=2)
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$phylogenetic_repulsion %>% filter(!parameter %in% c("discrepancy","Ynew","prediction","pred_error")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

Mean phylogenetic covariance martix

```{r}
tauC<-function(omega,gamma,I,D,lambda_cov){
      C = exp(-lambda_cov * D)
      vCov=omega*C + (1-omega) * I
      result<-solve(vCov*gamma)
      result<-reshape2::melt(result)
      return(result)
}

result<-chains$phylogenetic_repulsion %>% filter(parameter %in% c("omega","gamma","lambda_cov")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(tauC(omega=.$omega,gamma=.$gamma,I=I,D=D,lambda_cov=.$lambda_cov)) %>% group_by(Var1,Var2) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(result,aes(x=Var1,y=Var2,fill=mean)) + geom_tile() + scale_fill_gradient2(low="red",mid="white",high="blue")
```

## Decay in phylogenetic repulsion

```{r}

effect_repulsion<-function(lambda,D,omega,gamma){
  C<-exp(-lambda*D)
  vCov=(omega*C[,] + (1-omega) * I)
  vCov=solve(vCov)*gamma
  data.frame(lambda,Distance=as.numeric(D),Covariance=as.numeric(vCov))
}

phyrepulse_effect<-chains$phylogenetic_repulsion %>% filter(parameter %in% c("lambda_cov","omega","gamma")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(effect_repulsion(lambda=.$lambda_cov,D=D,omega=.$omega,gamma=.$gamma)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(phyrepulse_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```

# Traits

## Trait Attraction

```{r}
#Distance matrix
rownames(traits)<-traits[,1]
Dtraits<-as.matrix(dist(traits[,-1]))
Dtraits<-Dtraits/max(Dtraits)

#filter data to match traits
indat<-transects_zero %>% filter(Plant %in% rownames(traits))  %>%  mutate(n=(n>0)*1)

#Easiest to work with jags as numeric ordinal values
indat$Plant<-as.factor(indat$Plant)
indat$jPlant<-as.numeric(indat$Plant)

jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=1:length(levels(indat$Plant)))

#smmoth data
indat<-indat %>% group_by(Plant,jPlant,Site,Month,Year) %>% summarize(n=max(n))

#hold 2017 data aside for prediction
to_predict<-indat %>% filter(Year %in% "2017")
indat<-indat %>% filter(!Year %in% "2017")

#Add index
indat$Index<-as.factor(1:nrow(indat))
to_predict$Index<-as.factor(1:nrow(to_predict))
```

```{r}
models$trait_attraction<-runModel_autocorrelation(Yobs_dat=indat,Ynew_dat=to_predict,D=Dtraits,mod_name = "threshold_attraction")
```

#Get Chains

```{r}
chains$trait_attraction<-getChains(models$trait_attraction,indat,to_predict) 
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$trait_attraction %>% filter(parameter %in% c("alpha")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$trait_attraction %>% filter(par %in% c("fit","fitnew","lambda_cov","fitpred","omega","gamma")) %>%  ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=2)
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$trait_attraction %>% filter(!parameter %in% c("discrepancy","Ynew","prediction","pred_error")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

```{r}
tauC<-function(omega,gamma,I,D,lambda_cov){
      C = exp(-lambda_cov * D)
      vCov=omega*C + (1-omega) * I
      result<-vCov*gamma
      result<-reshape2::melt(result)
      return(result)
}

result<-chains$trait_attraction %>% filter(parameter %in% c("omega","gamma","lambda_cov")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(tauC(omega=.$omega,gamma=.$gamma,I=I,D=D,lambda_cov=.$lambda_cov)) %>% group_by(Var1,Var2) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(result,aes(x=Var1,y=Var2,fill=mean)) + geom_tile() + scale_fill_gradient2(low="red",mid="white",high="blue")
```

## Decay in trait attraction

```{r}
trait_effect<-chains$trait_attraction %>%  filter(parameter %in% c("lambda_cov","omega","gamma")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(effect(lambda=.$lambda_cov,D=D,omega=.$omega,gamma=.$gamma)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))
ggplot(trait_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```

## Repulsion

```{r}
models$trait_repulsion<-runModel_autocorrelation(Yobs_dat=indat,Ynew_dat=to_predict,D=Dtraits,mod_name = "threshold_repulsion")
```

#Get Chains

```{r}
chains$trait_repulsion<-getChains(models$trait_repulsion,indat,to_predict) 
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$trait_repulsion %>% filter(parameter %in% c("alpha")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$trait_repulsion %>% filter(par %in% c("fit","fitnew","lambda_cov","omega","fitpred","gamma")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=2)
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$trait_repulsion %>% filter(!parameter %in% c("discrepancy","Ynew","prediction","pred_error")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

```{r}
tauC<-function(omega,gamma,I,D,lambda_cov){
      C = exp(-lambda_cov * D)
      vCov=omega*C + (1-omega) * I
      result<-solve(vCov*gamma)
      result<-reshape2::melt(result)
      return(result)
}

result<-chains$trait_repulsion %>% filter(parameter %in% c("omega","gamma","lambda_cov")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(tauC(omega=.$omega,gamma=.$gamma,I=I,D=D,lambda_cov=.$lambda_cov)) %>% group_by(Var1,Var2) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(result,aes(x=Var1,y=Var2,fill=mean)) + geom_tile() + scale_fill_gradient2(low="red",mid="white",high="blue")
```

## Decay in trait repulsion

```{r}
traitrepulse_effect<-chains$trait_repulsion %>% filter(parameter %in% c("lambda_cov","omega","gamma")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(effect_repulsion(lambda=.$lambda_cov,D=D,omega=.$omega,gamma=.$gamma)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(traitrepulse_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```


# Interaction

## Attraction

```{r,fig.height=15}
#prepare data
indat<-transects_zero %>% filter(Plant %in% rownames(Dint))  %>%  mutate(n=(n>0)*1)

#Easiest to work with jags as numeric ordinal values
indat$Plant<-as.factor(indat$Plant)
indat$jPlant<-as.numeric(indat$Plant)


jagsIndexPlants<-data.frame(Plant=levels(indat$Plant),jPlant=1:length(levels(indat$Plant)))

indat<-indat %>% group_by(Plant,jPlant,Site,Month,Year) %>% summarize(n=max(n))

#hold 2017 data aside for prediction
to_predict<-indat %>% filter(Year %in% "2017")
indat<-indat %>% filter(!Year %in% "2017")

#Add index
indat$Index<-as.factor(1:nrow(indat))
to_predict$Index<-as.factor(1:nrow(to_predict))
```

```{r}
models$interaction_attraction<-runModel_autocorrelation(Yobs_dat=indat,Ynew_dat=to_predict,D=Dint,mod_name = "threshold_repulsion")
```

```{r}
chains$interaction_attraction<-getChains(models$interaction_attraction,indat,to_predict)
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$interaction_attraction %>% filter(parameter %in% c("alpha","e")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$interaction_attraction %>% filter(parameter %in% c("fit","fitnew","lambda_cov","omega","fitpred","gamma")) %>%  ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=2)
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$interaction_attraction %>% filter(!parameter %in% c("discrepancy","Ynew","prediction","pred_error")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

Mean interaction covariance

```{r}
tauC<-function(omega,gamma,I,D,lambda_cov){
      C = exp(-lambda_cov * D)
      vCov=omega*C + (1-omega) * I
      result<-vCov*gamma
      result<-reshape2::melt(result)
      return(result)
}

result<-chains$interaction_attraction %>% filter(parameter %in% c("omega","gamma","lambda_cov")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(tauC(omega=.$omega,gamma=.$gamma,I=I,D=Dint,lambda_cov=.$lambda_cov)) %>% group_by(Var1,Var2) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(result,aes(x=Var1,y=Var2,fill=mean)) + geom_tile() + scale_fill_continuous(low="white",high="red")
```

## Decay in interaction attraction

```{r}

int_effect<-chains$interaction_attraction %>% filter(parameter %in% c("lambda_cov","omega","gamma")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(effect(lambda=.$lambda_cov,D=D,omega=.$omega,gamma=.$gamma)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(int_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")

```

## Repulsion

```{r}
models$interaction_repulsion<-runModel_autocorrelation(Yobs_dat=indat,Ynew_dat=to_predict,D=Dint,mod_name = "threshold_repulsion")
```

#Get Chains

```{r}
chains$interaction_repulsion<-getChains(models$interaction_repulsion,indat,to_predict) 
```

### Evaluate convergence

```{r,fig.height=12,fig.width=15}
chains$interaction_repulsion %>% filter(parameter %in% c("alpha","e")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_grid(parameter~Plant,scales="free")
```

```{r}
chains$interaction_repulsion %>% filter(par %in% c("fit","fitnew","lambda_cov","omega","fitpred","gamma")) %>% ggplot(.,aes(x=Draw,col=as.factor(chain),y=value)) + geom_line() + facet_wrap(~parameter,scales="free",ncol=2)
```

### Posterior estimates
```{r,fig.height=10,fig.width=10}
chains$interaction_repulsion %>% filter(!parameter %in% c("discrepancy","Ynew","prediction","pred_error")) %>% ggplot(.,aes(x=value,fill=Plant)) + geom_histogram() + facet_wrap(~parameter,ncol=2,scales="free")
```

Mean interaction covariance martix

```{r}
tauC<-function(omega,gamma,I,D,lambda_cov){
      C = exp(-lambda_cov * D)
      vCov=omega*C + (1-omega) * I
      result<-solve(vCov*gamma)
      result<-reshape2::melt(result)
      return(result)
}

result<-chains$interaction_repulsion %>% filter(parameter %in% c("omega","gamma","lambda_cov")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(tauC(omega=.$omega,gamma=.$gamma,I=I,D=Dint,lambda_cov=.$lambda_cov)) %>% group_by(Var1,Var2) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

ggplot(result,aes(x=Var1,y=Var2,fill=mean)) + geom_tile() + scale_fill_gradient2(low="red",mid="white",high="blue")
```

## Decay in interaction repulsion

```{r}

intrepulse_effect<-chains$interaction_repulsion %>% filter(parameter %in% c("lambda_cov","omega","gamma")) %>% select(-par) %>% spread(parameter,value) %>% group_by(Draw,chain) %>% do(effect_repulsion(lambda=.$lambda_cov,D=D,omega=.$omega,gamma=.$gamma)) %>% group_by(Distance) %>% filter(!Distance==0) %>% summarize(mean=mean(Covariance),lower=quantile(Covariance,0.05),upper=quantile(Covariance,0.95))

ggplot(intrepulse_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance")
```

# Model Comparison

```{r}
#reduce models to a single df
for(x in 1:length(chains)){
  chains[[x]]$Model<-names(chains)[[x]]
}

allchains<-bind_rows(chains)

#create an attraction/repulsion label
allchains$Effect<-str_detect(allchains$Model,"repulsion")

allchains[!allchains$Effect==TRUE,"Effect"]<-"Attraction"
allchains[allchains$Effect==TRUE,"Effect"]<-"Repulsion"
allchains[allchains$Model=="julian","Effect"]<-NA
```

## Alpha

The probability of occurrence.

```{r,fig.height=12,fig.width=10}
#Alphas

#The baseline model is not logit transformed intercept
alphas<-allchains %>% filter(parameter %in% "alpha")

alphas[!alphas$Model %in% "baseline","value"]<-inv.logit(alphas[!alphas$Model %in% "baseline","value"])

alphas %>% group_by(Model,Plant) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% ggplot(.,aes(x=Model,y=mean)) + geom_pointrange(aes(ymin=lower,ymax=upper)) + coord_flip() + facet_wrap(~Plant,scales="free")
```

## E: The effect of autocorrelation on mean flowering intensity

```{r,fig.height=12,fig.width=10}

allpars<-allchains %>% filter(!parameter %in% c("discrepancy")) %>% group_by(Model,Plant,parameter) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))
```


### E: The effect of autocorrelation


```{r}
allchains %>% filter(parameter %in% "e") %>% group_by(Model,jPlant,Site,Month) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% filter(Site==1) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=inv.logit(mean))) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="Effect on p(Occ)") 
```

# By site

## 1300m - 1500m

```{r}
allchains %>% filter(parameter %in% "Ynew") %>% group_by(Model,jPlant,Site,Month) %>% summarize(p=sum(value)/n()) %>% filter(Site==1) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=p)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="p(Occ)") 
```

```{r}
 allchains %>% filter(parameter %in% "discrepancy") %>% group_by(Model,jPlant,Site,Month) %>% summarize(mean=mean(value)) %>% filter(Site==1) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=mean)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="Discrepancy") 

```

## 1500-1700m

```{r}
allchains %>% filter(parameter %in% "Ynew") %>% group_by(Model,jPlant,Site,Month) %>% summarize(p=sum(value)/n()) %>% filter(Site==2) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=p)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="p(Occ)") 
```


```{r}
 allchains %>% filter(parameter %in% "discrepancy") %>% group_by(Model,jPlant,Site,Month) %>% summarize(mean=mean(value)) %>% filter(Site==2) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=mean)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="Discrepancy") 

```

## 1700-1900 

```{r}
allchains %>% filter(parameter %in% "Ynew") %>% group_by(Model,jPlant,Site,Month) %>% summarize(p=sum(value)/n()) %>% filter(Site==3) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=p)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="p(Occ)") 
```

```{r}
 allchains %>% filter(parameter %in% "discrepancy") %>% group_by(Model,jPlant,Site,Month) %>% summarize(mean=mean(value)) %>% filter(Site==3) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=mean)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="Discrepancy") 

```

## 1900-2100m

```{r}
allchains %>% filter(parameter %in% "Ynew") %>% group_by(Model,jPlant,Site,Month) %>% summarize(p=sum(value)/n()) %>% filter(Site==4) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=p)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="p(Occ)") 
```


```{r}
 allchains %>% filter(parameter %in% "discrepancy") %>% group_by(Model,jPlant,Site,Month) %>% summarize(mean=mean(value)) %>% filter(Site==4) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=mean)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="Discrepancy") 

```

## 2100-2300m

```{r}
allchains %>% filter(parameter %in% "Ynew") %>% group_by(Model,jPlant,Site,Month) %>% summarize(p=sum(value)/n()) %>% filter(Site==5) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=p)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="p(Occ)") 
```


```{r}
 allchains %>% filter(parameter %in% "discrepancy") %>% group_by(Model,jPlant,Site,Month) %>% summarize(mean=mean(value)) %>% filter(Site==5) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=mean)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="Discrepancy") 

```

## 2300m - 2500m

```{r}
allchains %>% filter(parameter %in% "Ynew") %>% group_by(Model,jPlant,Site,Month) %>% summarize(p=sum(value)/n()) %>% filter(Site==6) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=p)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="p(Occ)") 
```


```{r}
 allchains %>% filter(parameter %in% "discrepancy") %>% group_by(Model,jPlant,Site,Month) %>% summarize(mean=mean(value)) %>% filter(Site==6) %>% inner_join(jagsIndexPlants) %>% ggplot(.,aes(x=as.factor(Month),y=Plant)) + geom_tile(aes(fill=mean)) + facet_wrap(~Model) + scale_fill_continuous(low="blue",high="red") + labs(fill="Discrepancy") 

```
## Omega: The magnitude of the effect of autocorrelation on mean flowering occurrence

```{r}
allpars %>% filter(parameter %in% "omega") %>% ggplot(.,aes(x=Model,y=mean)) + geom_pointrange(aes(ymin=lower,ymax=upper)) + coord_flip() + labs(y="Omega")
```

## Gamma: The variance of the effect of autocorrelation on mean flowering occurrence

```{r,fig.height=12,fig.width=10}
allpars %>% filter(parameter %in% "gamma") %>% ggplot(.,aes(x=Model,y=mean)) + geom_pointrange(aes(ymin=lower,ymax=upper)) + coord_flip() + labs(y="Gamma")
```

## Decay in autocorrelation effect

```{r}
phy_effect$Model<-"Phylogenetic Attraction"
phyrepulse_effect$Model<-"Phylogenetic Repulsion"
trait_effect$Model<-"Trait Attraction"
traitrepulse_effect$Model<-"Trait Repulsion"
intrepulse_effect$Model<-"Interaction Repulsion"
int_effect$Model<-"Interaction Attraction"

est_effect<-bind_rows(list(phy_effect,phyrepulse_effect,trait_effect,traitrepulse_effect,traitrepulse_effect,int_effect,intrepulse_effect))
ggplot(est_effect,aes(x=Distance,y=mean)) + geom_line() + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.5) + theme_bw() + labs(x="Distance",y="Covariance") + facet_wrap(~Model,scales="free",ncol=2)
```

# Model Fit

## Bayesian pvalue

```{r,fig.height=10,fig.width=10}

fitstat<-allchains %>% filter(parameter %in% c("fit","fitnew"),!Model=="baseline") %>% select(Model,Effect,Draw,chain,parameter,value)  %>% spread(key=parameter,value=value) 

ymin<-min(fitstat$fit)
ymax<-max(fitstat$fit)
ab<-data.frame(x=seq(0,ymax,1),y=seq(0,ymax,1))

ggplot(data=fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data",col="Model")  + ggtitle("Posterior predictive check") + coord_fixed() + geom_abline()

#Bayesian p-value
fitstat %>% group_by(Model) %>% summarize(p=sum((fitnew>fit)/n()))
```

## Overall

```{r}
allchains %>% filter(parameter %in% "fit") %>% mutate(value=value) %>%  group_by(Model) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% ggplot(.,aes(y=mean,x=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Discrepancy") + theme_bw()

allchains %>% filter(parameter %in% "fit",!Model %in% "baseline") %>% mutate(value=value) %>%  group_by(Model) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% arrange(desc(mean)) %>% kable() %>% kable_styling()
```

### Without baseline

```{r}
allchains %>% filter(parameter %in% "fit",!Model %in% "baseline") %>% mutate(value=value) %>%  group_by(Model) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% ggplot(.,aes(y=mean,x=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Discrepancy") + theme_bw()
```

## By Species

```{r}
allchains %>% filter(parameter %in% "discrepancy") %>% group_by(Draw,chain,Model,Plant) %>% summarize(value=sqrt(sum(value)/n())) %>% group_by(Model,Plant) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% arrange(mean) %>% ggplot(.,aes(x=Plant,y=mean,col=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Discrepancy") + theme_bw()
```

Without baseline

```{r}
allchains %>% filter(parameter %in% "discrepancy",!Model=="baseline") %>% group_by(Draw,chain,Model,Plant) %>% summarize(value=sqrt(sum(value)/n())) %>% group_by(Model,Plant) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% arrange(mean) %>% ggplot(.,aes(x=Plant,y=mean,col=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Discrepancy") + theme_bw()
```

Zoom in 
```{r,fig.height=10,fig.width=12}
allchains %>% filter(parameter %in% "discrepancy",!Model=="baseline") %>% group_by(Draw,chain,Model,Plant) %>% summarize(value=sqrt(sum(value)/n())) %>% group_by(Model,Plant) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% arrange(mean) %>% ggplot(.,aes(y=mean,x=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + labs(y="Root Mean Squared Error") + theme_bw() + facet_wrap(~Plant,ncol=3,scales="free") + coord_flip()
```

# Prediction

```{r}
allchains %>% filter(parameter %in% "fitpred") %>% mutate(value=value) %>%  group_by(Model) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% ggplot(.,aes(y=mean,x=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Squared Error") + theme_bw()
```

```{r}
allchains %>% filter(parameter %in% "pred_error",!Model=="baseline") %>% group_by(Draw,chain,Model,Plant) %>% summarize(value=sqrt(sum(value)/n())) %>% group_by(Model,Plant) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% arrange(mean) %>% ggplot(.,aes(x=Plant,y=mean,col=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Discrepancy") + theme_bw()
```

```{r}
allchains %>% filter(parameter %in% "fitpred") %>% mutate(value=value) %>%  group_by(Model) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% ggplot(.,aes(y=mean,x=Model)) + geom_pointrange(aes(ymin=lower,ymax=upper),alpha=0.5,position=position_dodge(width=0.4)) + coord_flip() + labs(y="Squared Error") + theme_bw() 
```

### Tables

```{r}
allchains %>% filter(parameter %in% "fitpred") %>% mutate(value=value) %>%  group_by(Model) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95)) %>% arrange(desc(mean)) %>% kable() %>% kable_styling()
```

```{r}
save.image("/Users/ben/Dropbox/Gesner/Phenology.RData")
```
