---
title: "R Notebook"
output: html_notebook
---

#Phylogenetic model of species interactions

```{r}

#Interactions
interactions<-read.csv("data/cleaned/interactions.csv",row.names=1)
head(interactions) %>% kable() %>% kable_styling()
```

# Phylogeny

```{r,eval=FALSE}
phy<-read.newick("data/SantaLucia_tree_Newick")
species<-str_match(phy$tip.label,"_(\\w+)")[,2]
tspecies<-unique(interactions$Plant)
to_join<-data.frame(Plant=tspecies,phylo=str_match(tspecies," (\\w+)")[,2])
new_names<-data.frame(phylo=species) %>% inner_join(to_join)
phy$tip.label<- as.character(new_names$Plant)
plot(phy)

#variance covariance matrix
vcov<-vcv(phy,corr = T)

#Identity matrix
I<-diag(nrow(vcov))
```

# Traits
```{r}
traits<-read.csv("data/Species_SantaLucia.csv")

#improve column names
traits<-traits %>% select(Plant=Iplant_Double,min_corolla_tube=Lt,corolla_length=Lc,vertical_diameter= D,stamen_length=S,stamen_exertion=E)

#order the factors

head(traits) %>% kable() %>% kable_styling()
```

Sum interaction matrix plotted in plant phylogenetic order.

```{r}
m<-interactions %>% filter(Plant %in% phy$tip.label) %>% group_by(Hummingbird,Plant) %>% summarize(n=n())
m$Plant<-factor(m$Plant,levels=phy$tip.label)
ggplot(m,aes(x=Hummingbird,y=Plant,fill=n)) + geom_tile() + theme_bw() + theme(axis.text.x=element_text(angle=-90))
```

Do plants show phylogenetic signal in hummingbird partners?

Start with binary interactions

```{r}
#Source model
source("model/intercept_phylogeny.R")

#print model
writeLines(readLines("model/intercept_phylogeny.R"))

#Run Model
runModel<-function(Yobs_dat,I,vcov){
  
  
  #Parameters to track
  ParsStage <- c("alpha","lambda","sigma","e","tauC")
  
  #Jags Data


  Dat<-list(
    Yobs=Yobs_dat$Yobs,
    Birds=max(c(Yobs_dat$jBird)),
    Bird=Yobs_dat$jBird,
    Plant=Yobs_dat$jPlant,
    Plants=max(c(Yobs_dat$jPlant)),
    Nobs=length(Yobs_dat$Yobs),
    I=I,
    vcov=vcov,
    zeros=rep(0,max(c(Yobs_dat$jPlant))))
  
  #MCMC options
    system.time(
      model<-jags(data=Dat,parameters.to.save=ParsStage,model.file="model/intercept_phylogeny.jags",n.thin=1,n.iter=100,n.burnin=0,n.chains=2,DIC=F)
    )
    return(model)
}

models$intercept<-runModel(Yobs_dat=indat,I=I,vcov=vcov)
```

#Get Chains
```{r}
getChains<-function(mod){
  
  pc<-reshape2::melt(mod$BUGSoutput$sims.array)
  colnames(pc)<-c("Draw","chain","par","value")
  
  #extract parameter name
  pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]
  
  #Extract index
  splitpc<-split(pc,pc$parameter)
  
  #single index
  splitpc[c("alpha","tauC")]<-lapply(
    splitpc[c("alpha","tauC")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+)]"))[,3:4]
    colnames(sv)<-c("jBird","jPlant")
    pc<-data.frame(x,sv)
    })
  
    splitpc[c("e")]<-lapply(
    splitpc[c("e")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,jPlant=sv)
    })
    
  chains<-bind_rows(splitpc)
return(chains)}
```

## Alpha

The probability of interactions among hummingbirds and plants

Grab 1 species to visualize

```{r,fig.height=8,fig.width=8}
chains$intercept<-getChains(models$intercept)

#merge species names
chains$intercept<-chains$intercept %>% left_join(jagsIndexBird) %>% left_join(jagsIndexPlants)

#alpha
chains$intercept %>% filter(parameter %in% "alpha",jPlant %in% c(5,10)) %>% ggplot(.,aes(x=inv.logit(value),fill=Plant)) + geom_histogram() + facet_wrap(~Hummingbird,scales="free")
```

```{r,fig.height=8,fig.width=8}
chains$intercept %>% filter(parameter %in% "alpha") %>% group_by(Hummingbird,Plant) %>% summarize(n=mean(value)) %>% ggplot(.,aes(fill=inv.logit(n),x=Plant,y=Hummingbird)) + geom_tile() + scale_fill_continuous(low="blue",high="red") + theme(axis.text.x = element_text(angle=-90)) + labs(fill="P(Interaction)")
```

## Lambda

The phylogenetic signal in interaction probability

```{r}
chains$intercept %>% filter(parameter %in% "lambda") %>% ggplot(.,aes(x=Draw,y=value,col=as.factor(chain))) + geom_line() 

chains$intercept %>%  filter(parameter %in% "lambda") %>% ggplot(.,aes(x=value,fill=par)) + geom_histogram()

```

## sigma

The variance phylogenetic signal in interaction probability

```{r}
chains$intercept %>% filter(parameter %in% "sigma") %>% ggplot(.,aes(x=Draw,y=value,col=as.factor(chain))) + geom_line() 

chains$intercept %>%  filter(parameter %in% "sigma") %>% ggplot(.,aes(x=value,fill=par)) + geom_histogram()
```

# e

This can be thought of as the effect size of phylogenetic influence. Change in probability of interactions due to interactions

```{r}
chains$intercept %>%  filter(parameter %in% "e") %>% ggplot(.,aes(fill=Plant,x=inv.logit(value))) + geom_histogram() + labs(fill="Probability of interactions")
```


```{r}
save.image("/Users/ben/Dropbox/Gesner/analysis.RData")
```
